### 2014-10-01 exploring_sem/ggm_pacakge.Rmd
file changes

---
title: "Intro to the `ggm` R package for graphical Markov models"
output: html_document
---

From the package description:
>This package provides functions for defining, manipulating and fitting graphical Markov models.

For application in structural equation modeling and path analysis I am initially interested in the functionality for defining graphs. Particularly, I have a complex time-series path analysis diagramming a 1-year lag-effect. 
![My almost graph]("CH2_DAG.svg")
I want to define the graph and independence claims it outlines.
```{r}
library(ggm) #Load library

```