### 2014-12-02 analysis/exploring_sem/lavaan_tutorial.Rmd
revamped directory and made some changes to repeated measures and exploratory script files. plan to change these to Rmd

---
title: "Intro to `lavaan` pacakage for SEM - Tutorial"
output: html_document
---

#### Load `lavaan` package and create an SEM using appropriate syntax
```{r load lavaan, echo=TRUE}
library(lavaan)
myModel <- ' # regressions
             y1 + y2 ~ f1 + f2 + x1 + x2
                  f1 ~ f2 + f3
                  f2 ~ f3 + x1 + x2

             # latent variable definitions 
               f1 =~ y1 + y2 + y3 
               f2 =~ y4 + y5 + y6 
               f3 =~ y7 + y8 + y9 + y10

             # variances and covariances 
               y1 ~~ y1 
               y1 ~~ y2 
               f1 ~~ f2

             # intercepts 
               y1 ~ 1 
               f1 ~ 1
           '
```

#### Example 1: Fitting a confirmatory factor analysis (CFA) model
```{r CFA Model, echo=TRUE}
HS.model <- ' visual  =~ x1 + x2 + x3 
              textual =~ x4 + x5 + x6
              speed   =~ x7 + x8 + x9 '
fit <- cfa(HS.model, data = HolzingerSwineford1939)
summary(fit, fit.measures = TRUE) # Get summary of fitted model
```

#### Example 2: Fitting a structural equation model (SEM)
```{r SEM Model, echo=TRUE}
model <- '
  # measurement model (latent variable definitions, CFA)
    ind60 =~ x1 + x2 + x3
    dem60 =~ y1 + y2 + y3 + y4
    dem65 =~ y5 + y6 + y7 + y8

  # regressions
    dem60 ~ ind60
    dem65 ~ ind60 + dem60

  # residual correlations (covariances)
    y1 ~~ y5
    y2 ~~ y4 + y6
    y3 ~~ y7
    y4 ~~ y8
    y6 ~~ y8
'
fit <- sem(model, data = PoliticalDemocracy)
summary(fit, standardized = TRUE)
```