### 2015-05-06 search4superspreaders.Rmd
Making a presentation on data exploration to try and identify superspreader individuals

---
output:
  slidy_presentation:
    font_adjustment: 1
    footer: Center for Geospatial Analytics at NCSU
    highlight: haddock
    keep_md: yes
  beamer_presentation:
    highlight: haddock
---

## Searching for Superspreaders

**Whalen Dillon**

May 7, 2015

<img src=http://cnr.ncsu.edu/geospatial/wp-content/uploads/sites/6/2014/07/cgaDynamicB2REV.png width="800px" />

## Emerging Infectious Diseases

New in a population, or new but rapidly increasing

![](presentations_images/ebola_lyme_whitenose.png)
<img src=presentations_images/global_shipping.png width="500px" />

## Disease Triangle

![](presentations_images/disease_triangle.png)

## Sudden Oak Death Triangle

![](presentations_images/sod_triangle.png)

## Sudden Oak Death Hosts & Environment

46 wildland host species - 5 key host tree species

<img src=presentations_images/sod_host_distribution.png />

## Pathogen Spillover

**“when disease epidemics are driven by transmission from an alternate reservoir host population” – Power & Mitchell (2004)**
- characteristic of multihost disease systems

![](presentations_images/sod_spillover.png)

## Superspreaders

Individual(s) that disproportionately contribute to disease transmission
 - potential mechanism for pathogen spillover

![](presentations_images/bay_superspreader.png)

## Searching for superspreaders

```{r}
summary(cars)
```

## Slide with Plot

```{r, echo=FALSE}
plot(cars)
```

## Species and Locations as Superspreaders

Disease hotspots as "superspreaders"

![](presentations_images/superspreader_hotspots.png)