### 2015-05-08 search4superspreaders.Rmd
Created table of 90th, 95th, 99th quantile thresholds.  Plotted all 99th quantile stems, subsets of all that were in the 99th 2+ years, and all in the 99th in 3+ years. Ended presentation on hotspots as superspreaders.

---
output:
  slidy_presentation:
    fig_width: 8
    fig_height: 5.5
    font_adjustment: 1
    footer: Center for Geospatial Analytics at NCSU
    highlight: haddock
    incremental: yes
    keep_md: yes
  ioslides_presentation:
    highlight: haddock
    incremental: yes
  beamer_presentation:
    highlight: haddock
---

## Searching for Superspreaders

**Whalen Dillon**

May 7, 2015

<img src=http://cnr.ncsu.edu/geospatial/wp-content/uploads/sites/6/2014/07/cgaDynamicB2REV.png width="800px" align="center" />

## Emerging Infectious Diseases

New in a population, or new but rapidly increasing

![](presentations_images/ebola_lyme_whitenose.png)
<img src=presentations_images/global_shipping.png width="500px" />

## Disease Triangle

![](presentations_images/disease_triangle.png)

## Sudden Oak Death Disease Triangle

![](presentations_images/sod_triangle.png)

## Sudden Oak Death Hosts & Environment

46 wildland host species - 5 key host tree species

<img src=presentations_images/sod_host_distribution.png />

## Pathogen Spillover

**“when disease epidemics are driven by transmission from an alternate reservoir host population”**

-- Power & Mitchell (2004) *Pathogen spillover in disease epidemics*

![](presentations_images/sod_spillover.png)

A common characteristic of multihost disease systems

## Superspreaders

Individual(s) that disproportionately contribute to disease transmission

 -- potential mechanism for pathogen spillover

![](presentations_images/bay_superspreader.png)

```{r load data & libraries, echo=FALSE, message=FALSE}
load("~/GitHub/superspreaders/stem_level_data.RData")
library(plyr); library(dplyr); library(xtable)
```
```{r subset to stems w/slc, echo=FALSE}
slc <- bay_laurel %>% select(plot, tag, slc, date, status, year) %>% 
      filter(status == "Alive" & slc != "NA")
#summary(slc)
slc <- droplevels(slc)
```

Number of bay stems in the study alive during at least one sampling year: **`r length(unique(slc$tag))`**

## Searching for superspreaders

The 90%, 95%, and 99% quantile threshold values for each year
```{r Look at potential outliers, echo=FALSE, message=FALSE, results='asis', out.extra=''}
tab <- (bay_laurel %>% select(plot, tag, slc, date, status, year) %>% 
      filter(status == "Alive") %>%
      group_by(year) %>% 
      summarise(#slc80 = quantile(slc, 0.80, na.rm = T),
                #slc85 = quantile(slc, 0.85, na.rm = T), 
                slc90 = quantile(slc, 0.90, na.rm = T), 
                slc95 = quantile(slc, 0.95, na.rm = T),
                slc99 = quantile(slc, 0.99, na.rm = T)))
tab <- xtable(tab)
print(tab, type = "html")
```

## Searching for superspreaders
```{r Plot slc99 by year, message=FALSE, echo=FALSE}
library(ggplot2)
g <- ggplot(tab, aes(year, slc99))
g + geom_bar(stat = "identity", color = "blue", fill = "skyblue") + 
      theme_bw() + theme(legend.position = "none") +
      ggtitle("99th Percentiles of Symptomatic Leaf Counts by Year") +
      ylab("Symptomatic Leaf Count")
```

```{r Subset data to potential outliers each year, echo=FALSE}
slc90q <- bay_laurel %>% group_by(year) %>% 
      filter(year == 2004 & slc >= 82 | year == 2005 & slc >= 63 | 
                   year == 2006 & slc >= 112 | year == 2007 & slc >= 72 | 
                   year == 2008 & slc >= 62 | year == 2009 & slc >= 65 | 
                   year == 2010 & slc >= 104 | year == 2011 & slc >= 166 |
                   year == 2012 & slc >= 126 | year == 2014 & slc >= 93)
slc95q <- bay_laurel %>% group_by(year) %>% 
      filter(year == 2004 & slc >= 114 | year == 2005 & slc >= 88 | 
                   year == 2006 & slc >= 130 | year == 2007 & slc >= 86 | 
                   year == 2008 & slc >= 76 | year == 2009 & slc >= 83 | 
                   year == 2010 & slc >= 139 | year == 2011 & slc >= 190 |
                   year == 2012 & slc >= 155 | year == 2014 & slc >= 128)
slc99q <- bay_laurel %>% group_by(year) %>%
      filter(year == 2004 & slc >= 178 | year == 2005 & slc >= 127 | 
                   year == 2006 & slc >= 174 | year == 2007 & slc >= 113 | 
                   year == 2008 & slc >= 110 | year == 2009 & slc >= 119 | 
                   year == 2010 & slc >= 191 | year == 2011 & slc >= 214 |
                   year == 2012 & slc >= 200 | year == 2014 & slc >= 181)
```

## Searching for Superspreaders - Stems in the 99th Percentile
```{r subset 99th to 5 variables, echo=FALSE}
slc99q <- slc99q %>% select(plot, tag, year, date, slc)
```
Number of observations in 99th percentiles: **`r length(slc99q$tag)`**

```{r Plot slc99 data by year, echo=FALSE}
qplot(date, slc, data = slc99q, geom = "jitter", color = tag) + 
      geom_line(aes(color = tag)) +
      theme_bw() + theme(legend.position = "none") +
      ylab("Symptomatic Leaf Count") + 
      ggtitle("Symptomatic Leaf Counts of Stems in 99th Percentile Each Year")
slc99q <- ungroup(slc99q)
```

Number of unique stems in 99th percentiles: **`r length(unique(slc99q$tag))`**

## Stems in the 99th Percentile 2+ Years
```{r subset stems in 99th 2+ years,}
d <- subset(slc99q, duplicated(slc99q$tag) | duplicated(slc99q$tag, fromLast = T))
d <- arrange(d, tag, year)
d
```

## Stems in the 99th Percentile 2+ Years
```{r plot stems in 99th 2+ yrs, echo=FALSE}
qplot(date, slc, data = d, color = tag, geom = "jitter") + 
      geom_line(aes(color = tag)) +
      #geom_text(aes(label = tag)) +
      theme_bw() + theme(legend.position = "none") +
      ylab("Symptomatic Leaf Count") +
      ggtitle("Stems in the 99th Percentiles 2+ Years")
```

Number of unique stems in the 99th percentile 2+ years: **`r length(unique(d$tag))`**

## Stems in the 99th Percentile 3+ Years
```{r echo=FALSE}
dt <- table(d$tag)
d2 <- subset(d, tag %in% names(dt[dt > 2]))
```

Number of unique stems in the 99th percentile 3+ years: **`r length(unique(d2$tag))`**

```{r echo=FALSE, eval=FALSE}
#summary(d)
#str(d)
#d <- droplevels(d)
#length(unique(d$plot))
#length(unique(d$tag))

dt <- table(d$tag)
#dt > 2
#dt[dt > 2]
#names(dt[dt > 2])
#d$tag %in% names(dt[dt > 2])
d2 <- subset(d, tag %in% names(dt[dt > 2]))
#d2
length(unique(d2$tag))
```
```{r Plot stems 3+ yrs 99th, echo=FALSE}
qplot(date, slc, data = d2, color = tag:plot, geom = "jitter") + 
      geom_line(aes(color = tag:plot)) +
      #geom_text(aes(label = plot)) +
      theme_bw() +
      ylab("Symptomatic Leaf Count") +
      ggtitle("Stems in 99th Percentiles 3+ Years")
```

## Species and Locations as Superspreaders

More symptoms due to:

 - individual characteristics
 - environmental characteristics
 - pathogen characteristics

![](presentations_images/superspreader_hotspots.png)

Disease hotspots as "superspreaders"