### 2015-06-24 iale2015_sem.Rmd
Starting analysis for IALE presentation. Trying to develop a path analysis to relate environmental variability, especially weather, to pathogen spillover

---
title: "IALE 2015 Analysis"
author: "Whalen Dillon"
date: "June 15, 2015"
output: html_document
---

```{r load data}
# Ecological data: leaf counts, infection
load("~/GitHub/superspreaders/stems_plots.RData")

# Physical environment: topography
library(rgdal)
plots_env <- readOGR("analysis/data/plot_gis/", "soco_plots_metric")
str(plots_env)
plot(plots_env, axes=T)
plots_env <- plots_env@data

# Temperature: rainy season hour summaries
plots_temps <- read.csv("analysis/data/rainy_season_temps.csv")
summary(plots_temps)
plots_temps <- select(plots_temps, -X)
```

Join the plot level environmental data (geographic/topographic & temperatures)
```{r join plot data, message=FALSE}
library(dplyr)
plots_env <- left_join(plots_env, plots_temps, by = c("PLOT_ID"="plotid"))
summary(plots_env)
str(plots_env)
plots_env <- plots_env %>% select(-cat) %>% rename(easting = X, northing = Y, plotid = PLOT_ID)
```

Investigating `NAs` in the data set
```{r investigate NAs}
summary(plots_env)
filter(plots_env, is.na(sample_year))

# BUSH01 plot included. I think we have limited ecological data for this plot, and no microclimate data. So, I am going to exclude it from all the data sets
plots_env <- filter(plots_env, plotid != "BUSH01")
stems <- rename(stems, plotid = plot) %>% filter(plotid != "BUSH01")

summary(plots_env) # A number of observations for above 25 threshold have NAs
# I believe that these are in fact truly zero observations. NAs occurred for the 2010, 2011, and 2014 seasons, and the only NAs in the data set are associated with the above 25 threshold. 
plots_env %>% filter(is.na(hrs_abv25))
# I am replacing the NAs with `0`, assuming true observations.
plots_env[is.na(plots_env)] <- 0
```

Make a pairs plot of the continuous variables, i.e. everything except `plotid`.
```{r pairs plot}
# Helper functions for `pairs`
panel.hist <- function(x, ...)
{
      usr <- par("usr"); on.exit(par(usr))
      par(usr = c(usr[1:2], 0, 1.5) )
      h <- hist(x, plot = FALSE)
      breaks <- h$breaks; nB <- length(breaks)
      y <- h$counts; y <- y/max(y)
      rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
      usr <- par("usr"); on.exit(par(usr))
      par(usr = c(0, 1, 0, 1))
      r <- abs(cor(x, y, use = "na.or.complete"))
      txt <- format(c(r, 0.123456789), digits = digits)[1]
      txt <- paste0(prefix, txt)
      if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
      text(0.5, 0.5, txt, cex = cex.cor * r)
}

pairs(select(plots_env, -plotid, -sample_year), lower.panel = panel.cor, diag.panel = panel.hist)

plots_env$avg_psi <- rowMeans(cbind(plots_env$psiSpr, plots_env$psiSum, plots_env$psiWin))
pairs(select(plots_env, -plotid, -sample_year, -starts_with("psi")), lower.panel = panel.cor, diag.panel = panel.hist)
```