### 2015-06-27 explore_veg_data.Rmd
add new file where i explore the data from vegetation transects

---
title: "Explore Vegetation Data"
author: "Whalen Dillon"
date: "June 26, 2015"
output: html_document
---

In this document I am exploring vegetation data from the Sonoma County plot network. These data were collected usin vegetation transects across each of the plots circa plot establishment (2003/2004) and in 2011. The DBH of alternate and non-host species were also recorded in 2006, 2012, and 2014.

Read in general vegetation data related to each plot
```{r read in data}
canopy_cover <- read.csv("analysis/data/canopy_cover_qry.csv")
veg_transects <- rbind(read.csv("analysis/data/veg_transects_subset_qry_pre2011.csv"), read.csv("analysis/data/veg_transects_2011.csv"))
alt_sp_dbh <- read.csv("analysis/data/alt_species_dbh.csv")
```

What am I looking at here?
```{r examine canopy cover data}
library(dplyr)
str(canopy_cover)
head(canopy_cover)
# I am interested in the date and total estimated cover fore each plot
canopy_cover <- canopy_cover %>% select(PlotID, Date, PctTotalCover)
canopy_cover$Date <- as.Date(as.character(canopy_cover$Date), format = "%m/%d/%Y")
canopy_cover <- rename(canopy_cover, plotid = PlotID, date = Date, pct_cvr = PctTotalCover)
summary(canopy_cover)
filter(canopy_cover, is.na(date))
filter(canopy_cover, plotid == "APGAR01") # appears to be a duplicate entry...
canopy_cover <- filter(canopy_cover, is.na(date) == "FALSE")

library(lubridate)
canopy_cover$year <- year(canopy_cover$date)
```

The subset veg transect data `veg2` is approximately half of the original data points, selecting every other sample point along the transect. About twice as many points were sampled during plot establishment compared to the 2011 sampling
```{r examine veg transect data}
str(veg_transects)
veg_transects$Date <- as.Date(as.character(veg_transects$Date), format = "%m/%d/%Y")
veg_transects$year <- year(veg_transects$Date)
summary(veg_transects)
head(veg_transects, 20)
unique(veg_transects$year)

dim(filter(veg_transects, year != 2011))

unique(filter(veg_transects, year == 2005 | year == 2004 | year == 2003)$PlotID) # 205 plot IDs
unique(filter(veg_transects, year == 2011)$PlotID) # 199 plot IDs

```