### 2015-07-03 iale2015_sem.Rmd
Calculated plot level leaf count variables, joined plot-level environmental; calculated understory richness from vegetation transects, and overstory richness from tagged and untagged stem DBH data and saved this into a "richness" data frame. Calculated multiple DBH variables for tagged & untagged stems, combined those data frames together, and calculated relative density, relative dominance, and importance value for each species.

---
title: "IALE 2015 Analysis"
author: "Whalen Dillon"
date: "June 15, 2015"
output: html_document
---

```{r load data}
# Ecological data: leaf counts, infection
load("~/GitHub/superspreaders/stems_plots.RData")
# Vegetation data
veg_us_2005 <- read.csv("analysis/data/understory_veg_2005.csv")
veg_us_2011 <- read.csv("analysis/data/understory_veg_2011.csv")
# Untagged dbh data
untagged_dbh <- read.csv("analysis/data/untagged_dbh_summary.csv")

# Physical environment: topography
library(rgdal)
plots_shp <- readOGR("analysis/data/plot_gis/", "soco_plots_metric")
summary(plots_shp)
plot(plots_shp, axes=T)
plots_env <- plots_shp@data

# Temperature: rainy season hour summaries
library(plyr); library(dplyr)
plots_temps <- read.csv("analysis/data/rainy_season_temps.csv")
summary(plots_temps)
plots_temps <- select(plots_temps, -X)
```

Join the plot level environmental data (geographic/topographic & temperatures)
```{r join plot data, message=FALSE}
plots_env <- left_join(plots_env, plots_temps, by = c("PLOT_ID"="plotid"))
summary(plots_env)
str(plots_env)
plots_env <- plots_env %>% select(-cat) %>% rename(easting = X, northing = Y, plotid = PLOT_ID)
```

Investigating `NAs` in the `plots_env` data set
```{r investigate NAs}
summary(plots_env)
filter(plots_env, is.na(sample_year))

# BUSH01 plot included. I think we have limited ecological data for this plot, and no microclimate data. So, I am going to exclude it from all the data sets
plots_env <- filter(plots_env, plotid != "BUSH01")
stems <- rename(stems, plotid = plot) %>% filter(plotid != "BUSH01")

summary(plots_env) # A number of observations for above 25 threshold have NAs
# I believe that these are in fact truly zero observations. NAs occurred for the 2010, 2011, and 2014 seasons, and the only NAs in the data set are associated with the above 25 threshold. 
plots_env %>% filter(is.na(hrs_abv25))
# I am replacing the NAs with `0`, assuming true observations.
plots_env[is.na(plots_env)] <- 0
```

Examine the bay laurel data at individual and plot levels
```{r examine umca data}
summary(plots_umca)
# This only has number of infected, uninfected, and total, which must apply to living stems only. I want to analyze leaf counts.
summary(bay_laurel)
# This has stem-level data on bay laurel.
```

I can use this to calculate plot level leaf count variables (sum, average) for each plot and year
```{r calculate plot level leaf counts}
plots_umca <- left_join(plots_umca, bay_laurel %>% filter(status == "Alive") %>% 
      group_by(plot, year) %>% 
      summarise(
            avg_lfct = mean(slc, na.rm = T),
            tot_lfct = sum(slc, na.rm = T)),
      by = c("plot", "year"))
summary(plots_umca)
```

```{r examine lfct NAs}
filter(plots_umca, is.na(avg_lfct))
# I'm not sure why this wasn't excluded from the calculations with the na.rm = T option
filter(bay_laurel, plot == "YAHNG02", status == "Alive")
# I'm removing it manually from the `plots_umca` data frame
plots_umca <- plots_umca %>% filter(avg_lfct != "NaN")
summary(plots_umca)
```

So now I have a plot-level of UMCA infection levels for each year. I want to join this with some of the environmental data.

```{r join env and umca data}
summary(plots_env)
plots_env$avg_psi <- rowMeans(cbind(plots_env$psiSpr, plots_env$psiSum, plots_env$psiWin))
summary(plots_umca)
# plot == plotid, year == sample_year
plots_umca <- left_join(select(plots_umca, -species), plots_env, by = c("plot" = "plotid", "year" = "sample_year"))
summary(plots_umca)
filter(plots_umca, is.na(easting))
# NAs are due to no microclimate data for year 2003, and abandoned plots BUSH01 & SUGAR02
```

Now I also want to calculate some plot-level community variables from the vegetation data. First, understory richness from the vegetation transects.
```{r calculate understory richness}
summary(veg_us_2005)
veg_us_2005 <- as.tbl(select(veg_us_2005, -X))
summary(veg_us_2011)
veg_us_2011 <- as.tbl(select(veg_us_2011, -X))
veg_us_2011
veg_us_2005

veg <- rbind(veg_us_2005, veg_us_2011)
veg
unique(veg$year)
veg$year[veg$year < 2011] <- 2005 # Creates the two most complete veg sampling
veg <- veg %>% rename(us_species = UnderstorySp, plotid = PlotID)
richness <- veg %>% group_by(plotid, year) %>% summarise(us_rich = length(unique(us_species)))
richness
```

Next, overstory richness from the DBH data for tagged & untagged stems.

```{r calculate overstory richness}
summary(untagged_dbh)
untagged_dbh <- as.tbl(untagged_dbh %>% select(-X) %>% rename(plotid = PlotID, species = Species))
untagged_dbh
unique(untagged_dbh$year)
untagged_dbh$year[untagged_dbh$year < 2014] <- 2005 # Create two most complete groups
untagged_dbh$plotid <- tolower(untagged_dbh$plotid)

untagged_dbh %>% group_by(plotid, year) %>% summarise(untagged_rich = length(unique(species)))

stems <- as.tbl(stems)
summary(stems)
unique(stems$status)
stems <- stems %>% select(-X) %>% filter(status == "Alive" | status == "Dead")
unique(stems$species)
stems$species <- tolower(stems$species)
stems$species[stems$species == "quke x quag"] <- "quag x quke"
stems$plotid <- tolower(stems$plotid)
stems %>% group_by(plotid, year) %>% summarise(tagged_rich = length(unique(species)))

os_richness <- left_join(
      stems %>% group_by(plotid, year) %>% summarise(tagged_rich = length(unique(species))),
      untagged_dbh %>% group_by(plotid, year) %>% summarise(untagged_rich = length(unique(species))), by = c("plotid", "year")
) %>% mutate(os_rich = tagged_rich + untagged_rich)
os_richness
richness <- left_join(os_richness, richness, by = c("plotid","year"))
richness
summary(richness)
```

I also want to calculate some variables related to DBH, and in particular the "importance value" for each species. Importance value is calculated in the literature as:

`IV = 0.5 * (relative density / relative dominance)` where 
`Relative Density = # stems of species / total # of stems` and
`Relative Dominance = basal area of species / total basal area`

```{r calculate importance value}
unique(untagged_dbh$species)
untagged_dbh
unique(stems$species)

tagged_dbh <- left_join(stems %>% 
      filter(status == "Alive", location == "In") %>%
      select(plotid, year, species, dbh) %>% 
      group_by(plotid, species, year) %>% 
      summarise(live_count = length(species), live_avg_dbh = mean(dbh), live_tot_dbh = sum(dbh)),
      stems %>% 
      filter(status == "Dead", location == "In") %>%
      select(plotid, year, species, dbh) %>% 
      group_by(plotid, species, year) %>% 
      summarise(dead_count = length(species), dead_avg_dbh = mean(dbh), dead_tot_dbh = sum(dbh)),
      by = c("plotid", "species", "year")
)
tagged_dbh <- as.tbl(tagged_dbh)
summary(tagged_dbh)

tagged_dbh
untagged_dbh

stems_dbh <- rbind(tagged_dbh, untagged_dbh)

stems_dbh <- stems_dbh %>% group_by(plotid, year) %>% mutate(all_live_count = sum(live_count), all_live_dbh = sum(live_tot_dbh, na.rm = T))
summary(stems_dbh)
filter(stems_dbh, all_live_dbh > 800)

stems_dbh <- stems_dbh %>% mutate(rel_dens = live_count / all_live_count, rel_dom = live_tot_dbh / all_live_dbh, imp_value = 0.5*(rel_dens + rel_dom))
summary(stems_dbh)
```

Make a pairs plot of the continuous variables for UMCA plots.
```{r pairs plot}
# Helper functions for `pairs`
panel.hist <- function(x, ...)
{
      usr <- par("usr"); on.exit(par(usr))
      par(usr = c(usr[1:2], 0, 1.5) )
      h <- hist(x, plot = FALSE)
      breaks <- h$breaks; nB <- length(breaks)
      y <- h$counts; y <- y/max(y)
      rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
      usr <- par("usr"); on.exit(par(usr))
      par(usr = c(0, 1, 0, 1))
      r <- abs(cor(x, y, use = "na.or.complete"))
      txt <- format(c(r, 0.123456789), digits = digits)[1]
      txt <- paste0(prefix, txt)
      if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
      text(0.5, 0.5, txt, cex = cex.cor * r)
}

plots_umca <- ungroup(plots_umca) %>% select(-starts_with("psi"), -plot, -year)

pairs(select(plots_umca, -plot, -year, -starts_with("psi")), lower.panel = panel.cor, diag.panel = panel.hist)
```