### 2015-12-21 oak_infection_model.Rmd
Adjust data and models.

---
title: "Predicting Oak Infection"
author: "Whalen Dillon"
date: "June 11, 2015"
output: html_document
---

I'm trying to define an empirical relationship between oak infection and bay laurel density, though I think this is something we can get from Sarah's survival analysis.

```{r load data}
load("stems_plots.RData")
summary(stems)
summary(qusp_plots)
qusp_plots$plot <- tolower(qusp_plots$plot)
qusp_plots <- rename(qusp_plots, plotid = plot)
summary(umca_plots)
```

```{r load packages, message=FALSE}
library(plyr); library(dplyr); library(lme4); library(ggplot2)
```

```{r manipulate data}
names(stems)

inf_oak_stems <- stems %>% 
      select(plotid, year, species, tag, status, canker, dbh2) %>% 
      filter(species != "umca", species != "lide", species != "unknown sp.", canker != 0)

summary(inf_oak_stems)
summary(filter(inf_oak_stems, year == 2011))
filter(inf_oak_stems, year == 2014)
qplot(dbh2, data = inf_oak_stems)

plot_aggregates <- stems %>%
      select(plotid, slc, canker, sod_dead, dbh, year) %>%
      group_by(year, plotid) %>% 
      summarise(slc_total = sum(slc, na.rm=T), slc_avg = mean(slc, na.rm=T), oak_inf = sum(canker, na.rm=T), oak_dead = sum(sod_dead, na.rm=T), dbh_total = sum(dbh, na.rm=T), dbh_avg = mean(dbh, na.rm=T))

summary(plot_aggregates)

summary(qusp_plots)
plot_aggregates <- left_join(plot_aggregates, select(qusp_plots, year, plotid, uninfected_oak_ct, infected_oak_ct, tot_oak), by = c("year", "plotid"))
summary(umca_plots)
plot_aggregates <- left_join(plot_aggregates, select(umca_plots, year, plotid, uninfected_bay_ct, infected_bay_ct, tot_bay))

plot_aggregates
filter(plot_aggregates, year == 2011) %>% group_by(plotid) %>% summarise(inf_oak_ct = sum(infected_oak_ct, na.rm = T))

plot_aggregates %>% group_by(year) %>% summarise(oak_inf = sum(oak_inf), oak_dead = sum(oak_dead))
sum(plot_aggregates$oak_inf, na.rm = T)
```

```{r model oak infection}
library(lme4)
oak2012 <- filter(plot_aggregates, year == 2012, infected_oak_ct > 0)
summary(oak2012)
plot(oak2012$tot_bay, oak2012$infected_oak_ct)

summary(filter(plot_aggregates, infected_oak_ct == 1))

ggplot(plot_aggregates, aes(tot_bay, infected_oak_ct)) +
      geom_point() +
      facet_grid(year ~ .)
ggplot(plot_aggregates)

oak.m1 <- glmer(cbind(infected_oak_ct, uninfected_oak_ct) ~ I(tot_bay/225) + (1 | plotid), data = plot_aggregates, family = "binomial")
summary(oak.m1)

oak.m2 <- glmer(cbind(infected_oak_ct, uninfected_oak_ct) ~ log1p(tot_bay) + (1 | plotid), data = plot_aggregates, family = "binomial")
summary(oak.m2)
```