### 2015-12-21 vegan_diversity.Rmd
Add script exploring diversity analysis using 'vegan' package.

---
title: "vegan Diversity"
author: "Whalen Dillon"
date: "August 24, 2015"
output: html_document
---

```{r load vegan package, echo=FALSE, results='hide'}
library(vegan)
data(BCI)
summary(BCI)
str(BCI)
head(BCI)
names(BCI)
```

The data example is a wide format, with each species a column and the number (abundunce) of that species observed in each row. Each row corresponds to the observation unit, in this case 1-hectare plots.

The `diversity` function finds the Shannon-Wiener, Simpson, and inverse Simpson diversity indices:

  - Shannon-Wiener: $H = -\sum_{i = 1}^{S} {(p_{i}) (log_b p_i)}$
  - Simpson: $D_1 = 1 - \sum_{i = 1}^{S} p_i^2$ 
  - inverse Simpson: $D_2 = 1 / \sum_{i = 1}^{S} p_i^2$

     - $p_i$ = proportion of species *i*
     - *S* = number of species
     - *b* = base of logarithm

The default uses the natural logarithm.

```{r calculate diversity}
H <- diversity(BCI, index = "shannon", MARGIN = 1, base = exp(1))
D1 <- diversity(BCI, index = "simpson")
D2 <- diversity(BCI, index = "invsimpson")
H
D1
D2
```

#### Species Abundance Models

Fisher's log-series estimates the expected number of species with *n* individuals, plotting species by frequencies. 
```{r Fishers log series}
k <- sample(nrow(BCI), 1)
fish <- fisherfit(BCI[k,])
plot(fish)
fish
```

Preston's log-normal model bins species into frequency classes of increasing sizes instead of plotting species by frequencies. There are two alternative fucntions in the `vegan` pacakge for fitting the lognormal model:

  1. `prestonfit` uses binning approach with arbitrary choices of limits and ties
     - for ties between adjacent octaves half were in 1st octave and half transferred to next octave
     - can either split ties or keep all limit cases in lower octave
  2. `prestondistr` maximizes truncated log-normal likelihood without binning data
     - recommended in `vegan` vignette paper

```{r Prestons lognormal}
prest <- prestondistr(BCI[k,])
plot(prest)
prest
```

This uses maximum-likelihood method and, I think, log base 2 (?)

#### Whittaker Plots - Ranked Abundance Distribution
The `vegan` function `radfit` uses maximum likelihood to fit popular models:

  - brokenstick: null model with no estimated parameters in `vegan`
  - preemption: one estimated parameter ($\alpha$)
  - log-normal: two estimated parameters ($\mu, sigma$)
  - Zipf: two estimated parameters (p1, $\gamma$)
  - Zipf-Mandlebrot: three estimated parameters (*c*, $\beta, gamma$)

The models are customarily defined for proportions, but `radfit` works directly with abundance data.

```{r Whittaker plots}
rad <- radfit(BCI[k,], family = poisson)
plot(rad)
radlattice(rad)
rad
```

This function compares models using AIC or BIC (based on log-likelihood, penalized by number of estimated parameters).