### 2016-02-15 veg-change-assessment.Rmd
Add veg change assessment document.

---
title: "Vegetation Change Assessment - SOD Plots in Sonoma County"
author: "Whalen Dillon"
date: "February 15, 2016"
output: html_document

---

```{r}
boxplot(tot_ba_m2 ~ year, 
        data = basal_abund %>% filter(status == "Alive", species == "UMCA"),
        ylab = "m^2",
        main = "Live UMCA Basal Area")
boxplot(tot_ba_m2 ~ year, 
        data = basal_abund %>% filter(status == "Dead", species == "UMCA"),
        ylab = "m^2",
        main = "Dead UMCA Basal Area")

boxplot(tot_ba_m2 ~ year, 
        data = basal_abund %>% filter(status == "Alive", species == "QUAG"),
        ylab = "m^2",
        main = "Live QUAG Basal Area")
boxplot(tot_ba_m2 ~ year, 
        data = basal_abund %>% filter(status == "Dead", species == "QUAG"),
        ylab = "m^2",
        main = "Dead QUAG Basal Area")

ggplot(data = basal_abund %>% filter(status == "Alive", species == "QUAG"),
       aes(year, tot_ba_m2, group = year)) +
      geom_boxplot()

boxplot(abundance ~ year, 
        data = basal_abund %>% filter(status == "Alive", species == "QUAG"),
        ylab = "Number of stems",
        main = "Live QUAG Abundance")
boxplot(tot_ba_m2 ~ year, 
        data = basal_abund %>% filter(status == "Dead", species == "QUAG"),
        ylab = "Number of stems",
        main = "Dead QUAG Abundance")

ggplot(data = basal_abund,
       aes(species, tot_ba_m2, label = plotid)) +
      geom_boxplot() +
      facet_grid(status ~ year) +
      geom_text() +
      ylab("Basal Area (m^2)")

filter(dbhs, plotid == "SUGAR05", species == "QUAG")
# new giant dead stem in 2014, was probably dead entire study period

ggplot(data = basal_abund,
       aes(species, abundance)) +
      geom_boxplot() +
      facet_grid(status ~ year) +
      #geom_text() +
      ylab("Abundance")

ggplot(basal_abund %>% filter(status == "Alive"), aes(tot_ba_m2)) + 
      geom_histogram() +
      facet_grid(species ~ year)
ggplot(basal_abund %>% filter(status == "Alive"), aes(avg_ba_m2)) + 
      geom_histogram() +
      facet_grid(species ~ year)
ggplot(basal_abund %>% filter(status == "Alive"), aes(abundance)) + 
      geom_histogram() +
      facet_grid(species ~ year)
```