### 2016-02-16 veg-change-assessment.Rmd
Update script.
Added read-in of data and initial examination of overstory/understory species data. Created some plots of species abundance: number of plots a species was recorded in, and the total number of a species recorded across the entire plot network. These are split by year, plot establishment (2005) and the most recent sampling season (2014).

---
title: "Vegetation Change Assessment - SOD Plots in Sonoma County"
author: "Whalen Dillon"
date: "February 15, 2016"
output: 
  html_document:
    toc: true
---

```{r read in data, echo=F, results='hide', message=FALSE}
library(plyr, verbose = F); library(dplyr, verbose = F);
library(ggplot2, verbose = F); library(tidyr, verbose = F)

# Tagged Stems
stems <- as.tbl(read.csv("analysis/data/stems_noBUSH01.csv"))
stems
stems$plotid <- tolower(stems$plotid)
stems$species <- tolower(stems$species)

# Overstory species data of tagged and untagged stems
tag_dbh_plot <- as.tbl(read.csv("analysis/data/tag-dbh-plot.csv"))
tag_dbh_plot
tag_dbh_plot$plotid <- tolower(tag_dbh_plot$plotid)
tag_dbh_plot$species <- tolower(tag_dbh_plot$species)
tag_dbh_stem <- as.tbl(read.csv("analysis/data/tag-dbh_0514-corrected.csv"))
tag_dbh_stem
tag_dbh_stem$plotid <- tolower(tag_dbh_stem$plotid)
tag_dbh_stem$species <- tolower(tag_dbh_stem$species)
alt_dbh_plot <- as.tbl(read.csv("analysis/data/alt-dbh-plot.csv"))
alt_dbh_plot
alt_dbh_stem <- as.tbl(read.csv("analysis/data/alt-species-dbh-corrected.csv"))
alt_dbh_stem

unique(alt_dbh_plot$plotid) %in% unique(tag_dbh_plot$plotid)
unique(alt_dbh_plot$plotid)[c(25,47,61,74,112,174,180)]
# ann28: abandoned early b/c no hosts, remove
# badge01: no hosts, kept in study
# bush01, halow01, sweet01: abandoned early, remove
# mcneil01, votru01: not sampled in 2014, remove
alt_dbh_plot <- filter(alt_dbh_plot, plotid != "ann28", plotid != "bush01", plotid != "halow01", plotid != "sweet01", plotid != "mcneil01", plotid != "votru01")
os_dbh_plot <- rbind(
      select(tag_dbh_plot, plotid, species, status, year, everything()), 
      select(alt_dbh_plot, -tot_dbh, -avg_dbh))
sort(unique(os_dbh_plot$species))

# Understory species data from microplots of veg transects
veg_sub <- as.tbl(read.csv("analysis/data/us-veg-sub.csv"))
veg_sub
sort(unique(veg_sub$us_species))

# Understory and overstory diversity/evenness calculations
richness <- as.tbl(read.csv("analysis/data/os_us_diversity.csv"))
richness

# Importance Values
imp_val_plot <- as.tbl(read.csv("analysis/data/species_dbh_impval_plot.csv"))
imp_val_plot
imp_val_plot <- select(imp_val_plot, -X)
summary(imp_val_plot)
```

```{r number of species, echo=F}
length(unique(os_dbh_plot$species))
length(unique(veg_sub$us_species))
ggplot(veg_sub %>% group_by(plotid, year, us_species), 
       aes(us_species)) +
      geom_bar()

# Get total number of plots each understory species occurred in at least once
ggplot(veg_sub %>% group_by(year, us_species) %>% 
             summarise(abundance = length(plotid)), 
       aes(us_species, abundance)) +
      geom_bar(stat = "identity") +
      facet_grid(year ~.)

# Plot total number of plots each overstory species occurred in at least once
ggplot(os_dbh_plot %>% filter(status == "Alive") %>% 
             group_by(year, species) %>% 
             summarise(abundance = length(plotid)), 
       aes(species, abundance)) +
      geom_bar(stat = "identity") +
      facet_grid(year ~.)
```

```{r number of overstory species}
# Plot total number of living stems for each overstory species by year
ggplot(os_dbh_plot %>% filter(status == "Alive") %>% 
             group_by(year, species), 
       aes(species, abundance)) +
      geom_bar(stat = "identity") +
      facet_grid(year ~.)

os_abund <- os_dbh_plot %>% 
      filter(status == "Alive") %>%
      group_by(year, species) %>% 
      summarise(abundance = sum(abundance)) %>% 
      spread(species, abundance)

## same as above, but filtering out the super-abundant UMCA
ggplot(os_dbh_plot %>% filter(status == "Alive", species != "umca") %>% 
             group_by(year, species), 
       aes(species, abundance)) +
      geom_bar(stat = "identity") +
      facet_grid(year ~.)
```

```{r plotting, echo=F}
ggplot(data = os_dbh_plot,
       aes(species, tot_ba_m2, label = plotid)) +
      geom_boxplot() +
      facet_grid(status ~ year) +
      #geom_text() +
      ylab("Basal Area (m^2)")

filter(dbhs, plotid == "SUGAR05", species == "QUAG")
# new giant dead stem in 2014, was probably dead entire study period

ggplot(data = basal_abund,
       aes(species, abundance)) +
      geom_boxplot() +
      facet_grid(status ~ year) +
      #geom_text() +
      ylab("Abundance")

ggplot(basal_abund %>% filter(status == "Alive"), aes(tot_ba_m2)) + 
      geom_histogram() +
      facet_grid(species ~ year)
ggplot(basal_abund %>% filter(status == "Alive"), aes(avg_ba_m2)) + 
      geom_histogram() +
      facet_grid(species ~ year)
ggplot(basal_abund %>% filter(status == "Alive"), aes(abundance)) + 
      geom_histogram() +
      facet_grid(species ~ year)

boxplot(tot_ba_m2 ~ year, 
        data = basal_abund %>% filter(status == "Alive", species == "UMCA"),
        ylab = "m^2",
        main = "Live UMCA Basal Area")
boxplot(tot_ba_m2 ~ year, 
        data = basal_abund %>% filter(status == "Dead", species == "UMCA"),
        ylab = "m^2",
        main = "Dead UMCA Basal Area")

boxplot(tot_ba_m2 ~ year, 
        data = basal_abund %>% filter(status == "Alive", species == "QUAG"),
        ylab = "m^2",
        main = "Live QUAG Basal Area")
boxplot(tot_ba_m2 ~ year, 
        data = basal_abund %>% filter(status == "Dead", species == "QUAG"),
        ylab = "m^2",
        main = "Dead QUAG Basal Area")

ggplot(data = basal_abund %>% filter(status == "Alive", species == "QUAG"),
       aes(year, tot_ba_m2, group = year)) +
      geom_boxplot()

boxplot(abundance ~ year, 
        data = basal_abund %>% filter(status == "Alive", species == "QUAG"),
        ylab = "Number of stems",
        main = "Live QUAG Abundance")
boxplot(tot_ba_m2 ~ year, 
        data = basal_abund %>% filter(status == "Dead", species == "QUAG"),
        ylab = "Number of stems",
        main = "Dead QUAG Abundance")
```