### 2016-03-22 veg-change-assessment.Rmd
Update variable names.

---
title: "Vegetation Change Assessment - SOD Plots in Sonoma County"
author: "Whalen Dillon"
date: "February 15, 2016"
output: 
  html_document:
    toc: true
---

## Load Packages and Data
```{r load packages, echo=F, results='hide', message=FALSE}
library(plyr, verbose = F); library(dplyr, verbose = F);
library(ggplot2, verbose = F); library(tidyr, verbose = F)
```

```{r load overstory tree data, echo=F, results='hide', message=FALSE}
# Tagged Stems
tagged_stems <- as.tbl(read.csv("analysis/data/tagged-stems-corrected.csv"))
tagged_stems
tagged_stems$plotid <- tolower(tagged_stems$plotid)
tagged_stems$species <- tolower(tagged_stems$species)

# Overstory species data of tagged and untagged stems
tag_dbh_plot <- as.tbl(read.csv("analysis/data/tag-dbh-plot.csv"))
tag_dbh_plot
tag_dbh_plot$plotid <- tolower(tag_dbh_plot$plotid)
tag_dbh_plot$species <- tolower(tag_dbh_plot$species)
tag_dbh_plot <- rename(tag_dbh_plot, stem_status = status)

tag_dbh_stem <- as.tbl(read.csv("analysis/data/tag-dbh_0514-corrected.csv"))
tag_dbh_stem
tag_dbh_stem$plotid <- tolower(tag_dbh_stem$plotid)
tag_dbh_stem$species <- tolower(tag_dbh_stem$species)
tag_dbh_stem <- rename(tag_dbh_stem, stem_status = status)
tag_dbh_stem <- left_join(tag_dbh_stem, select(tagged_stems, plotid, cluster, species, tag, location, stem_status, year, sod_killed))

alt_dbh_plot <- as.tbl(read.csv("analysis/data/alt-dbh-plot.csv"))
alt_dbh_plot
alt_dbh_plot <- rename(alt_dbh_plot, stem_status = status)

alt_dbh_stem <- as.tbl(read.csv("analysis/data/alt-species-dbh-corrected.csv"))
alt_dbh_stem
alt_dbh_stem <- rename(alt_dbh_stem, stem_status = status)

all_stems <- as.tbl(read.csv("analysis/data/all_stems_corrected.csv"))
summary(all_stems)
      # rbind(select(tag_dbh_stem, plotid, species, dbh, status, year),
      #             alt_dbh_stem)
all_stems$plotid <- as.factor(all_stems$plotid)
all_stems$stem_status <- as.factor(all_stems$stem_status)
all_stems$year <- as.factor(all_stems$year)

unique(alt_dbh_plot$plotid) %in% unique(tag_dbh_plot$plotid)
unique(alt_dbh_plot$plotid)[c(25,47,61,74,112,174,180)]
# ann28: abandoned early b/c no hosts, remove
# badge01: no hosts, kept in study
# bush01, halow01, sweet01: abandoned early, remove
# mcneil01, votru01: not sampled in 2014, remove
alt_dbh_plot <- filter(alt_dbh_plot, plotid != "ann28", plotid != "bush01", plotid != "halow01", plotid != "sweet01", plotid != "mcneil01", plotid != "votru01")
```

```{r recalculate plot data, echo=F}
# Recalculate tag-dbh-plot to exclude new large dead quag stem in 2014 at sugar05
filter(tag_dbh_stem, plotid == "sugar05", year == 2014)
tag_dbh_plot <- tag_dbh_stem %>% 
      filter(dbh >= 2.0, tag != 7794) %>% 
      group_by(plotid, species, year, stem_status, location) %>% 
      summarise(avg_ba_m2 = mean(pi*dbh^2/40000, na.rm = T), 
                tot_ba_m2 = sum(pi*dbh^2/40000, na.rm = T),
                abundance = length(dbh))
tag_dbh_plot <- ungroup(tag_dbh_plot)

# Overstory species w/measured DBH rooted in the plot
os_dbh_in_plot <- rbind(
      tag_dbh_plot %>% filter(location == "In") %>% 
            select(plotid, species, stem_status, year, everything(), -location), 
      select(alt_dbh_plot, -tot_dbh, -avg_dbh))
sort(unique(os_dbh_in_plot$species))

# Overstory species w/measured DBH in or overhanging the plot
os_dbh_plot <- rbind(
      tag_dbh_plot %>% group_by(plotid, species, stem_status, year) %>% 
            summarise(tot_ba_m2 = sum(tot_ba_m2), avg_ba_m2 = sum(tot_ba_m2)/sum(abundance),
                      abundance = sum(abundance)), 
      select(alt_dbh_plot, -tot_dbh, -avg_dbh))
```

```{r recalculate importance values, echo=F}
# Importance Values
# imp_val_plot <- as.tbl(read.csv("analysis/data/species_dbh_impval_plot.csv"))
# imp_val_plot# importance value by plot-species-year
# imp_val_plot <- select(imp_val_plot, -X)
# summary(imp_val_plot)
# arrange(select(imp_val_plot, plotid, species, year, live.ct = live_count, all.live.ct = all_live_count, live.tot.dbh = live_tot_dbh, all.live.dbh = all_live_dbh, imp.val = imp_value) %>% filter(year == 2014), plotid)
# unique(imp_val_plot$plotid)

impt_vals <- rbind(tag_dbh_stem %>% filter(location == "In", dbh >= 2.0) %>% 
                         select(plotid, species, dbh, stem_status, year), 
                   alt_dbh_stem) %>% 
      group_by(plotid, species, stem_status, year) %>% 
      summarise(sp_abund = length(species), 
                sp_basal_area = sum(sum(pi*dbh^2/40000, na.rm = T))) %>% 
      group_by(plotid, stem_status, year) %>% 
      mutate(total_abund = sum(sp_abund), 
             total_basal_area = sum(sp_basal_area), 
             imp_val =  
                   0.5 * ( (sp_abund / total_abund) + 
                                 (sp_basal_area / total_basal_area) ) )
summary(impt_vals)
impt_vals$plotid <- as.factor(impt_vals$plotid)
impt_vals$species <- as.factor(impt_vals$species)
impt_vals$year <- as.factor(impt_vals$year)
unique(impt_vals$plotid)
```

```{r load understory veg data, echo=F, results='hide', message=FALSE}
# Understory species data from microplots of veg transects
veg_us <- as.tbl(read.csv("analysis/data/us-veg.csv"))
veg_us
sort(unique(veg_us$us_species))
veg_sub <- as.tbl(read.csv("analysis/data/us-veg-sub.csv"))
veg_sub
sort(unique(veg_sub$us_species))
```

```{r load diversity metrics os and us}
# Understory and overstory diversity/evenness metrics
## These are based on the full set of understory species (n=77) and "overstory" species with DBH >= 2.0cm (tagged hosts) or DBH >= 5.0cm (untagged stems).
richness <- as.tbl(read.csv("analysis/data/os_us_diversity.csv"))
richness
```

## Visualize Species Occurrence and Abundance
We collected data on overstory and understory woody plant species to varying degrees throughout the study period. Overstory host species were assessed every year, so recruitment of new stems or death of existing stems was recorded, however, DBH measurements were initially only recorded at the first instance of a tagged host stem entering the study (DBH ≥ 2.0cm). We measured the DBH of untagged stems rooted in the plot and all tagged stems again in 2012. At this time we switched from an annual to bi-annual sampling cycle and decided that each time we visited the plots we would measure the DBH of all the overstory stems. Thus, the most recent DBH measurements are from the 2014 sampling season.

The understory species data was collected at two time points, 2005 and 2011. In 2005, Xnumber transects with Ynumber microplots each... Prior to doing the understory vegetation assessment in 2011 we conducted an analysis to assess the effort necessary to sufficiently collect comparable data. We repeatedly randomly sampled 50% of the entire dataset from 2005 and compared univariate statistics between this and the full dataset. Correlations were very strong (r > 0.9) and rare species were seldom missing from the smaller dataset.

During the 2011 season we used five transects across each plot with 0.5m radius microplots spaced at 1.3m intervals resulting in ~60 microplots per plot (12 per transect). Baselines were run between posts placed at plot plot establishment to clearly demarcate the plot edges. In some cases the distance between baselines was slightly greater or less than 15m, resulting in more or total number of microplots. At each microplot we recorded the occurrence of woody plant species that were rooted in or overhanging the microplot, and did not meet DBH requirements.

Initially there were 80 "species" in the understory vegetation data set as exported from the database, however, this list included two "unknowns" and "dead wood." We grouped some species to the genus level due to low confidence in species level identification, e.g. *Arctostaphylus* and removed the "dead wood" observation, reducing this to `r length(sort(unique(veg_us$us_species)))`.

### Understory Species
```{r number of understory species, echo=F}
length(sort(unique(veg_us$us_species)))
length(sort(unique(veg_sub$us_species)))

# Total number of plots each understory species occurred in at least once
ggplot(veg_us %>% group_by(year, us_species) %>% 
             summarise(occurrence = length(plotid)), 
       aes(us_species, occurrence, label = occurrence)) +
      geom_bar(stat = "identity") +
      geom_text(nudge_y = 3) +
      facet_grid(year ~ .) +
      ylab("# of plots") +
      ggtitle("Understory Species Occurrence (all species)")
ggplot(veg_us %>% group_by(year, us_species) %>% 
             summarise(occurrence = length(plotid)) %>% 
             filter(occurrence > 1), 
       aes(us_species, occurrence, label = occurrence)) +
      geom_bar(stat = "identity") +
      geom_text(nudge_y = 3) +
      facet_grid(year ~ .) +
      ylab("# of plots") +
      ggtitle("Understory Species Occurrence (2+ plots)")
length(
      unique(
            (veg_us %>% group_by(year, us_species) %>% 
             summarise(abundance = length(plotid)) %>% 
             filter(abundance > 1))$us_species
             )
       )
```

### Overstory Species
```{r number of overstory species rooted in the plot}
length(sort(unique(os_dbh_in_plot$species)))
summary(os_dbh_in_plot)
filter(os_dbh_in_plot, plotid == "sugar05")# dropped large, long-dead QUAG first recorded in 2014

# Total number of plots each overstory species occurred in at least once based on stems rooted in the plot
ggplot(os_dbh_in_plot %>% 
             filter(species != "unk1", species != "unknown sp.") %>% 
             group_by(year, species, stem_status) %>% 
             summarise(occurrence = length(plotid)), 
       aes(species, occurrence, label = occurrence)) +
      geom_bar(stat = "identity") +
      geom_text(nudge_y = 3) +
      facet_grid(stem_status ~ year) +
      ylab("# of plots") +
      ggtitle("Overstory Species Occurrence")

# Total number of living stems for each overstory species by year across all plots
ggplot(os_dbh_in_plot %>%
             filter(species != "unk1", species != "unknown sp.") %>% 
             group_by(year, species, stem_status) %>% 
             summarise(abundance = sum(abundance)), 
       aes(species, abundance, label = abundance)) +
      geom_bar(stat = "identity") +
      facet_grid(stem_status ~ year) +
      ylab("abundance") +
      ggtitle("Overstory Species Abundance") +
      geom_text(aes(vjust = 0), nudge_y = 10)

## same as above, but filtering out the super-abundant UMCA
ggplot(os_dbh_in_plot %>% 
             filter(species != "umca", species != "unk1", 
                    species != "unknown sp.") %>% 
             group_by(year, species, stem_status) %>% 
             summarise(abundance = sum(abundance)), 
       aes(species, abundance, label = abundance)) +
      geom_bar(stat = "identity") +
      facet_grid(stem_status ~ year) +
      ylab("abundance") +
      ggtitle("Overstory Species Abundance (minus UMCA)") +
      geom_text(aes(vjust = 0), nudge_y = 6)

## select overstory species abundance, abundance ≥ 100 stems
ggplot(os_dbh_in_plot %>%
             filter(species == "umca" | species == "quag" | species == "quga" | species == "psme" | species == "arme" | species == "quke" | species == "quch") %>% 
             group_by(year, species, stem_status) %>% 
             summarise(abundance = sum(abundance)), 
       aes(species, abundance, label = abundance)) +
      geom_bar(stat = "identity") +
      facet_grid(stem_status ~ year, scales = "free") +
      ylab("abundance") +
      ggtitle("Overstory Species Abundance") +
      geom_text(aes(vjust = 0), nudge_y = 8) +
      theme_bw()

ggplot(os_dbh_in_plot %>%
             filter(species == "umca" | species == "quag" | species == "quga" | species == "psme" | species == "arme" | species == "quke" | species == "quch" | species == "sese") %>% 
             group_by(year, species, stem_status) %>% 
             summarise(tot_ba_m2 = round(sum(tot_ba_m2), 2)), 
       aes(species, tot_ba_m2, label = tot_ba_m2)) +
      geom_bar(stat = "identity") +
      facet_grid(stem_status ~ year, scales = "free") +
      ylab("basal area") +
      ggtitle("Overstory Species Basal Area") +
      geom_text(aes(vjust = 0), nudge_y = 1.5) +
      theme_bw()

# Total numbers only for tagged host species. I think for 'Dead' status it is probably only fair to look at tagged host species because I don't think the dead stems of untagged species were frequently recorded earlier in the study.
## abundance
ggplot(os_dbh_in_plot %>% 
             filter(species == "lide" | species == "quag" | species == "quch" |
                          species == "quke" | species == "umca") %>%
             group_by(species, stem_status, year) %>% 
             summarise(abundance = sum(abundance)),
       aes(species, abundance, label = abundance)) +
      geom_bar(stat = "identity") +
      facet_grid(stem_status ~ year) +
      ylab("abundance") +
      ggtitle("Tagged Overstory Species Abundance") +
      geom_text(aes(vjust = 0), nudge_y = 6)

qusp_dbh_plot <- tag_dbh_stem %>% 
      filter(species == "quke" | species == "quag" | species == "quch") %>%
            group_by(plotid, species, stem_status, year, sod_killed) %>% 
            summarise(avg_ba_m2 = mean(pi*dbh^2/40000, na.rm = T), 
                tot_ba_m2 = sum(pi*dbh^2/40000, na.rm = T),
                abundance = length(dbh))
qusp_dbh_plot <- ungroup(qusp_dbh_plot)
summary(qusp_dbh_plot)
filter(qusp_dbh_plot, is.na(sod_killed))
qusp_dbh_plot$sod_killed[is.na(qusp_dbh_plot$sod_killed)] <- "No"

ggplot(qusp_dbh_plot, aes(species, abundance, label = abundance)) +
      geom_bar(stat = "identity") +
      facet_grid(sod_killed ~ year, scales = "free") +
      ylab("abundance") +
      ggtitle("Quercus Species SOD Killed and Living Abundance") #+
      geom_text(nudge_y = 6)
```

#### Total basal area of tagged species rooted in the plot
```{r basal area tagged species}
ggplot(os_dbh_in_plot %>% 
             filter(species == "lide" | species == "quag" | species == "quch" |
                          species == "quke" | species == "quag x quwi" | 
                          species == "quke x quag" | species == "umca") %>%
             group_by(species, stem_status, year) %>% 
             summarise(basal_area = round(sum(tot_ba_m2), digits = 4)),
       aes(species, basal_area, label = basal_area)) +
      geom_bar(stat = "identity") +
      facet_grid(stem_status ~ year) +
      ylab("basal area (m^2)") +
      ggtitle("Tagged Overstory Species Basal Area") +
      geom_text(aes(vjust = 0), nudge_y = 3)

# boxplot of total basal area for tagged species
ggplot(os_dbh_in_plot %>% 
             filter(species == "lide" | species == "quag" | species == "quch" |
                          species == "quke" | species == "quag x quwi" | 
                          species == "quke x quag" | species == "umca"),
       aes(species, tot_ba_m2, label = plotid)) +
      geom_boxplot() +
      facet_grid(stem_status ~ year) +
      ylab("basal area (m^2)") +
      ggtitle("Tagged Overstory Species Basal Area") +
      geom_text(nudge_y = .2)

# Sudden oak death mortality for QUAG, QUKE, QUCH
tag_dbh_stem$sod_dead[tag_dbh_stem$sod_dead==0] <- "alive"
tag_dbh_stem$sod_dead[tag_dbh_stem$sod_dead==1] <- "killed"
summary(tag_dbh_stem %>% 
      filter(species == "quke" | species == "quag" | species == "quch"))
qusp_dbh_plot <- tag_dbh_stem %>% 
      filter(species == "quke" | species == "quag" | species == "quch") %>%
            group_by(plotid, species, stem_status, year, sod_dead) %>% 
            summarise(avg_ba_m2 = mean(pi*dbh^2/40000, na.rm = T), 
                tot_ba_m2 = sum(pi*dbh^2/40000, na.rm = T),
                abundance = length(dbh))

ggplot(qusp_dbh_plot, aes(species, abundance, label = plotid)) +
      geom_boxplot() +
      facet_grid(sod_dead ~ year) +
      ylab("# of stems") +
      ggtitle("SOD Killed and Living Abundances (QUAG, QUKE, QUCH") #+
      geom_text(nudge_y = .2)

ggplot(qusp_dbh_plot, aes(species, tot_ba_m2, label = plotid)) +
      geom_boxplot() +
      facet_grid(sod_dead ~ year) +
      ylab("basal area (m^2)") +
      ggtitle("SOD Killed and Living Basal Area (QUAG, QUKE, QUCH") +
      geom_text(nudge_y = .2)
```

#### Species Importance Values
```{r importance values}
impt_vals <- ungroup(impt_vals)
unique(impt_vals$species)

ggplot(impt_vals %>% 
             filter(species == "lide" | species == "quag" | species == "quch" |
                          species == "quke" | species == "quag x quwi" | 
                          species == "quke x quag" | species == "umca"),
       aes(species, imp_val, label = plotid)) +
      geom_boxplot() +
      facet_grid(stem_status ~ year) +
      ylab("importance value") +
      ggtitle("Tagged Overstory Species Importance Value") #+
      geom_text(nudge_y = .2)
```

#### Species Richness
```{r diversity plots}
shannons_os <- richness %>% gather(key = year, value = diversity, -plotid) %>% 
      filter(year == "H.2005" | year == "H.2014")
shannons_os$year <- as.factor(ifelse(shannons_os$year == "H.2005", 2005, 2014))
summary(shannons_os)
pielous_os <- richness %>% gather(key = year, value = evenness, -plotid) %>% 
      filter(year == "J.2005" | year == "J.2014")
pielous_os$year <- as.factor(ifelse(pielous_os$year == "J.2005", 2005, 2014))

ggplot(shannons_os, aes(year, diversity, label = plotid)) +
      geom_boxplot() +
      # facet_grid(stem_status ~ year) +
      ylab("Shannon's Diversity Index") +
      ggtitle("Overstory Species Diversity") #+
      geom_text(nudge_y = .2)

ggplot(pielous_os, aes(year, evenness, label = plotid)) +
      geom_boxplot() +
      # facet_grid(stem_status ~ year) +
      ylab("Pielou's Evenness Index") +
      ggtitle("Overstory Species Evenness") #+
      geom_text(nudge_y = .2)
```
No difference in diversity across the entire plot network. No difference in average evenness, but evenness does appear to decrease for a subset of plots? Or, these plots are simply identified as outliers in 2014, but not in 2005. I am going to look more closely at those plots with an evenness index < 0.25 in the next chunk.

```{r evenness less than 0.25}
filter(pielous_os, evenness < 0.25, year == 2005)
filter(pielous_os, evenness < 0.25, year == 2014)

# plots that only occur in the 2005 data
anti_join(filter(pielous_os, evenness < 0.25, year == 2005),
          filter(pielous_os, evenness < 0.25, year == 2014),
          by = "plotid")
filter(pielous_os, plotid == "votru01", year == 2014)# plot decommisioned in 2011
filter(pielous_os, plotid == "cook02", year == 2014)# suspicious change in evenness, 0 to 0.34
filter(pielous_os, plotid == "backm03", year == 2014)# evenness increased from 0.19 to 0.33
filter(pielous_os, plotid == "athea02", year == 2014)# suspicious change in evenness, 0 to 0.62

# plots that only occur in the 2014 data
anti_join(filter(pielous_os, evenness < 0.25, year == 2014),
          filter(pielous_os, evenness < 0.25, year == 2005),
          by = "plotid")
filter(pielous_os, plotid == "pfend02", year == 2005)# decrease from 0.43 to 0.23
filter(pielous_os, plotid == "jroth03", year == 2005)# decrease from 0.57 to 0
filter(pielous_os, plotid == "jlsp05", year == 2005)# decrease from 0.43 to 0
filter(pielous_os, plotid == "hood02", year == 2005)# decrease from 0.30 to 0.25
filter(pielous_os, plotid == "ann22", year == 2005)# decrease from 0.46 to 0
```

The large changes toward or away from zero values for evenness are suspicious, and warrant investigating the original data. It seems unlikely that we would go from a single overstory species to multiple overstory species, and vice-versa, from multiple overstory species to a single overstory species.

```{r stems in low evenness plots}
summary(dbh_stem)
unique(filter(dbh_stem, plotid == "ann01")$species)
## plot dominated by umca, psme also present
unique(filter(dbh_stem, plotid == "ann01", year == 2005)$species)
## only umca in 2005
filter(dbh_stem, plotid == "ann01", year == 2014, species == "psme")
## single very large psme recorded in 2014, so it should have been recorded at some point earlier; first obs 2012, so made linear estimate for 2005 based on these two points and added to database.
```

#### Other Plots
```{r plotting, echo=F}
ggplot(data = os_dbh_in_plot,
       aes(species, tot_ba_m2, label = plotid)) +
      geom_boxplot() +
      facet_grid(stem_status ~ year) +
      #geom_text() +
      ylab("Basal Area (m^2)")

filter(dbhs, plotid == "SUGAR05", species == "QUAG")
# new giant dead stem in 2014, was probably dead entire study period

ggplot(data = os_dbh_in_plot,
       aes(species, abundance)) +
      geom_boxplot() +
      facet_grid(stem_status ~ year) +
      #geom_text() +
      ylab("Abundance")

ggplot(os_dbh_in_plot %>% filter(stem_status == "Alive"), aes(tot_ba_m2)) + 
      geom_histogram() +
      facet_grid(species ~ year)
ggplot(os_dbh_in_plot %>% filter(stem_status == "Alive"), aes(avg_ba_m2)) + 
      geom_histogram() +
      facet_grid(species ~ year)
ggplot(os_dbh_in_plot %>% filter(stem_status == "Alive"), aes(abundance)) + 
      geom_histogram() +
      facet_grid(species ~ year)

boxplot(tot_ba_m2 ~ year, 
        data = os_dbh_in_plot %>% filter(stem_status == "Alive", species == "UMCA"),
        ylab = "m^2",
        main = "Live UMCA Basal Area")
boxplot(tot_ba_m2 ~ year, 
        data = os_dbh_in_plot %>% filter(stem_status == "Dead", species == "UMCA"),
        ylab = "m^2",
        main = "Dead UMCA Basal Area")

boxplot(tot_ba_m2 ~ year, 
        data = os_dbh_in_plot %>% filter(stem_status == "Alive", species == "QUAG"),
        ylab = "m^2",
        main = "Live QUAG Basal Area")
boxplot(tot_ba_m2 ~ year, 
        data = os_dbh_in_plot %>% filter(stem_status == "Dead", species == "QUAG"),
        ylab = "m^2",
        main = "Dead QUAG Basal Area")

ggplot(data = basal_abund %>% filter(stem_status == "Alive", species == "QUAG"),
       aes(year, tot_ba_m2, group = year)) +
      geom_boxplot()

boxplot(abundance ~ year, 
        data = basal_abund %>% filter(stem_status == "Alive", species == "QUAG"),
        ylab = "Number of stems",
        main = "Live QUAG Abundance")
boxplot(tot_ba_m2 ~ year, 
        data = basal_abund %>% filter(stem_status == "Dead", species == "QUAG"),
        ylab = "Number of stems",
        main = "Dead QUAG Abundance")
```