### 2016-07-18 path-temp-compare.Rmd
Add residual plots for remaining models.
I also had to revert to the 1.1.1 version of the 'piecewiseSEM' package because of an error about matrices mis-matching when I used the most recent version. The error arises in the 'sem.missing.paths' function I think, but that's where my investigating ended.

---
title: "Path Model Temperature Comparisons"
author: "Whalen Dillon"
date: "April 14, 2016"
output: html_document
---

Based on previous studies, I expect that the summer season temperature and wet season conditions will influence leaf symptoms on California bay laurel. Specifically, hotter and drier summers will result in lower symptomatic leaf counts, because damaged leaves are more likely to abscise in these conditions. Increased abscision of infected leaves results in lower inoculum load to seed the seasonal epidemic during the following wet season. A wet season where rainfall extends into the late spring, when temperatures are warmer, will result in higher symptomatic leaf counts. These climate effects on California bay laurel leaf symptoms should manifest rapidly.

In contrast, symptoms on oak trees (bleeding or oozing cankers) are unlikely to be immediately detectable. Although infection of oak trees most likely happens when there is high inoculum load on California bay laurel, seldom will the infection be visibly detectable during the same year. I also assume that conditions need to be favorable for the infection to establish in the oak tree to a degree sufficient to cause visible symptoms. This is supported by Haas et al. (2016), who showed that oak trees experiencing two consecutive years of warm and wet conditions during November-May had an increased probability of exhibiting symptoms of infection. Since I am assessing infection based on visual symptoms I believe that the influence of conditions during the sampling season, particularly with respect to rainfall, are an effect on our ability to visually detect infection as opposed to a biological effect on infection of oak trees.

Based on these assessments of the differential relationships between our visual assessments of infection, and how climate and weather affect true infection, I developed temperature and moisture metrics to capture the effect on true infection. These metrics include dry-season temperature variables (the number of hours above or below a threshold temperature) that is intended to capture the influence on the amount of inoculum that over-summers on California bay laurel; a combined temperature/moisture variable that is the number of hours above, below, or within a temperature range during days that had measurable rainfall. Based on Haas et al. (2016), I expected the wet-hours variable averaged over the current and previous wet-seasons to have an effect on oak infection, but a limited statistical influence on California bay laurel leaf symptoms.

While temperature thresholds for growth, reproduction, and survival of *P. ramorum* have been experimentally assessed in laboratory settings, the conditions used have seldom corresponded to natural settings. For example, *P. ramorum* has a growth and reproduction optimum around 18-20C in laboratory experiments when humdity is 60-70% or greater. In the mediterranean climate of California rarely does a temperature above 18C correspond to humidity in that range. Still, these experiments and additional observations provided a starting point for developing a set of temperature thresholds to examine using a degree-day approach. By examining the relationships across multiple temperature cutoffs or ranges I am attempting to identify temperatures that may in fact be threshold temperatures where the inferential relationship between disease symptoms and climate variation changes.

```{r load data, echo=F}
load("climate-path-model-data.RData")
library(piecewiseSEM)# version 1.1.1, tried most recent version & kept getting errror 7/18/2016 b/c matrices rows were unequal in sem.missing.paths function
library(dplyr)
library(lme4)
library(lmerTest)

# Calculate oak density variable as stems per hectare 
oakplots.sub$umca_density_ha <- (oakplots.sub$tot_bay / (15*15))*10000
oakplots.sub$umca_density_ha.log <- log(oakplots.sub$umca_density_ha+1)
```

```{r combine plots data, eval=F, echo=F}
soco_plots <- stems %>%
      select(plotid, species, year) %>%
      group_by(plotid, year) %>%
      # filter(species == "UMCA", status == "Alive") %>%
      summarise(stem_total = length(species))

soco_plots <- rename(soco_plots, sample_year = year)

soco_plots <- left_join(soco_plots, select(umca_plots, plotid:tot_lfct), by = c("plotid", "sample_year"))
soco_plots <- left_join(soco_plots, select(oakplots.sub, plotid:inf_oak_ct), by = c("plotid","sample_year"))
soco_plots <- left_join(soco_plots, plots_env, by = "plotid")
summary(soco_plots)
soco_plots <- select(soco_plots, -stem_total)

write.csv(soco_plots, "analysis/data/soco_plots.csv")
```

```{r rescale variables, echo=F}
## rescale variables
# oakplots.sub2$inf_bay.scl <- scale((oakplots.sub2$infected_bay_ct))
oakplots.sub$tot_bay.scl <- as.numeric(scale(oakplots.sub$tot_bay))
oakplots.sub$tot_bay.log.scl <- as.numeric(scale(oakplots.sub$tot_bay.log))
oakplots.sub$umca_density_ha.scl <- as.numeric(scale(oakplots.sub$umca_density_ha))
oakplots.sub$umca_density_ha.log.scl <- as.numeric(scale(oakplots.sub$umca_density_ha.log))
oakplots.sub$umca_basal_area.scl <- as.numeric(scale(oakplots.sub$umca_basal_area))
oakplots.sub$umca_ba.log.scl <- as.numeric(scale(oakplots.sub$umca_ba.log))
oakplots.sub$tot_lfct.scl <- as.numeric(scale(oakplots.sub$tot_lfct))
oakplots.sub$tot_lfct.log.scl <- as.numeric(scale(oakplots.sub$tot_lfct.log))
# oakplots.sub$avg_lfct.scl <- scale(oakplots.sub$avg_lfct)
oakplots.sub$hrs_abv25ds.scl <- as.numeric(scale(oakplots.sub$hrs_abv25ds))
oakplots.sub$dys_abv25ds.scl <- as.numeric(scale(oakplots.sub$dys_abv25ds))
oakplots.sub$avg_tmax_ds.scl <- as.numeric(scale(oakplots.sub$avg_tmax_ds))
oakplots.sub$avg_hrsblw14_wet_tminus1.scl <- as.numeric(scale(oakplots.sub$avg_hrsblw14_wet_tminus1))
oakplots.sub$hrs1422_wet_tminus1.log.scl <- as.numeric(scale(oakplots.sub$hrs1422_wet_tminus1.log))
# oakplots.sub$avg_hrs1422_wet_t_t1.scl <- scale(oakplots.sub$avg_hrs1422_wet_t_t1)
oakplots.sub$avg_hrs1422_wet_t_t1.log.scl <- as.numeric(scale(oakplots.sub$avg_hrs1422_wet_t_t1.log))

## static variables
oakplots.sub$candens15m.scl <- as.numeric(scale(oakplots.sub$candens15m))
oakplots.sub$shannons2005.scl <- as.numeric(scale(oakplots.sub$H.2005))
oakplots.sub$shannons2014.scl <- as.numeric(scale(oakplots.sub$H.2014))
oakplots.sub$twi15m.scl <- as.numeric(scale(oakplots.sub$twi15m))
oakplots.sub$twi15m.log.scl <- as.numeric(scale(oakplots.sub$twi15m.log))
```

This code chunk describes the general structure for the path model for application in the `piecewiseSEM` package methods.
```{r general path model structure, eval=F}
model_list <- list(
      m1 <- lm(Diversity ~ Topographic_Index, data = data),
      m2 <- lm(Bay_Laurel_Density ~ Topographic_Index, data = data),
      m3 <- lm(Symptomatic_Leaf_Count ~ Bay_Laurel_Density + Dry_Season_Temperature + Topographic_Index, data = data),
      # m6 <- lm(candens15m ~ tot_bay.log, data = oakplots.2007),
      m6 <- lm(Wet_Hours_Temperature ~ Bay_Laurel_Density, data = data),
      m4 <- lm(Dry_Season_Temperature ~ Bay_Laurel_Density + Topographic_Index, data = data),
      m5 <- glm(Oak_Disease_Prevalence ~ Symptomatic_Leaf_Count + Wet_Hours_Temperature + Diversity, family = binomial(link = "logit"), data = data)
)

# model_list <- list(
#       m1 <- lm(H.2005 ~ twi15m.log, data = oakplots.2007),
#       m2 <- lm(tot_bay.log ~ twi15m.log, data = oakplots.2007),
#       m3 <- lm(tot_lfct.log ~ tot_bay.log + hrs_abv25ds + twi15m.log, data = oakplots.2007),
#       # m6 <- lm(candens15m ~ tot_bay.log, data = oakplots.2007),
#       m6 <- lm(avg_hrs1422_wet_t_t1 ~ tot_bay.log, data = oakplots.2007),
#       m4 <- lm(hrs_abv25ds ~ candens15m + twi15m.log, data = oakplots.2007),
#       m5 <- glm(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1 + H.2005 + candens15m, family = binomial(link = "logit"), data = oakplots.2007)
# )
```

Path model with the continuous variables standardized using the default settings of the `scale` function, resulting in each having a mean of zero and standard deviation of one. Each variable is first  centered by subtracting its mean and then scaled by dividing by its standard deviation.
```{r hours >25 path model all oak plots scaled}
model_list1.scl <- list(
      shannons <- lm(shannons2005.scl ~ twi15m.log.scl, data = oakplots.sub),

      umca_density <- lmer(umca_density_ha.log.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),

      slc <- lmer(tot_lfct.log.scl ~ umca_density_ha.log.scl + hrs_abv25ds.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),

      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log.scl ~ umca_density_ha.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),

      hrs_abv25 <- lmer(hrs_abv25ds.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),

      oak_infection <- glmer(oak.inf ~ tot_lfct.log.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.sub)
)
```

```{r hours >25 path model all oak plots unscaled, echo=F}
model_list1 <- list(
      shannons <- lm(H.2005 ~ twi15m.log, data = oakplots.sub),

      umca_density <- lmer(tot_bay.log ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),

      slc <- lmer(tot_lfct.log ~ tot_bay.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),

      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ tot_bay.log + (1|plotid) + (1|sample_year), data = oakplots.sub),

      hrs_abv25 <- lmer(hrs_abv25ds ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),

      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.sub)
)
```

Examining the residual plots of each piece of the path model
```{r shannons residuals plot}
par(mfrow=c(2,2))
plot(shannons, main = "OLS regression of diversity ~ wetness index residuals")
par(mfrow=c(1,1))
plot(shannons$model$twi15m.log, resid(shannons),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "Diversity model residuals vs. wetness index predictor")
```

```{r bay laurel density residuals plot}
plot(umca_density, main = "LMM of bay laurel density residuals")
plot(umca_density@frame$twi15m.log, resid(umca_density),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "Bay laurel density model residuals vs. wetness index predictor")
```

```{r symptomatic leaf count residuals plot}
plot(slc, main = "LMM of symptomatic leaf count residuals")
plot(slc@frame$umca_density_ha.log, resid(slc),
     xlab = "bay laurel density",
     ylab = "model residuals",
     main = "SLC model residuals vs. bay laurel density predictor")
plot(slc@frame$hrs_abv25ds, resid(slc),
     xlab = "# dry-season hours >25 C",
     ylab = "model residuals",
     main = "SLC model residuals vs. dry-season hours >25 C predictor")
plot(slc@frame$avg_hrs1422_wet_t_t1.log,
     xlab = "avg. wet-hours 14-22C",
     ylab = "model residuals",
     main = "SLC model residuals vs. wet-hours predictor")
plot(slc@frame$H.2005, resid(slc),
     xlab = "Shannon's diversity index",
     ylab = "model residuals",
     main = "SLC model residuals vs. Shannon's diversity predictor")
plot(slc@frame$twi15m.log, resid(slc),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "SLC model residuals vs. wetness index predictor")
```

```{r wet-hours residuals plot}
plot(wet_hours_1422, main = "LMM of avg. # of hours 14-22 C current & prev. year residuals")
plot(wet_hours_1422@frame$umca_density_ha.log, resid(wet_hours_1422),
     xlab = "bay laurel density",
     ylab = "model residuals",
     main = "Wet-hours model residuals vs. bay laurel density predictor")
```

```{r dry-season hours residuals plot}
plot(hrs_abv25, main = "LMM of # of hours >25 C during prev. dry season residuals")
plot(hrs_abv25@frame$twi15m.log, resid(hrs_abv25),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "Dry-season hours model residuals vs. wetness index predictor")
```

```{r oak infection residuals plot}
plot(oak_infection, main = "Binomial GLMM of oak infection residuals")
plot(oak_infection@frame$tot_lfct.log, resid(oak_infection),
     xlab = "symptomatic leaf count",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. symptomatic leaf count predictor")
plot(oak_infection@frame$avg_hrs1422_wet_t_t1.log, resid(oak_infection),
     xlab = "avg. wet-hours 14-22C",
     ylab = "model residuals",
     main = "Oak infectin model residuals vs. wet-hours predictor")
plot(oak_infection@frame$H.2005, resid(oak_infection),
     xlab = "Shannon's diversity index",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. diversity predictor")
```

Model set reduced to plots that have oaks & at least 1 bay laurel
```{r path model with oak plots having >0 bay laurel}
oak_and_bay_plots <- filter(oakplots.sub, tot_bay > 0)
oak_and_bay_list1 <- list(
      shannons <- lm(H.2005 ~ twi15m.log, data = oak_and_bay_plots),
      #umca_density <- lm(tot_bay.log ~ twi15m.log, data = oak_and_bay_plots),
      slc <- lmer(tot_lfct.log ~ tot_bay.log + hrs_abv25ds + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oak_and_bay_plots),
      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ tot_bay.log + (1|plotid) + (1|sample_year), data = oak_and_bay_plots),
      hrs_abv25 <- lmer(hrs_abv25ds ~ tot_bay.log + twi15m.log + (1|plotid) + (1|sample_year), data = oak_and_bay_plots),
      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oak_and_bay_plots)
)
```

```{r fit and compare all oak plots vs >0 bay laurel oak plots}
# Comparing fits of the piecewise models using all oak plots vs. plots with oaks and at least one bay laurel
sem.fit(model_list1, oakplots.sub, 
        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
sem.fit(model_list1.scl, oakplots.sub,
          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))

# sem.fit(oak_and_bay_list1, oak_and_bay_plots, 
#         corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log",
#                         "tot_bay.log ~~ H.2005"))
```

Get the path coeffcients for the fitted path models
```{r get path coefficients}
sem.coefs(model_list1, oakplots.sub,
          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
sem.coefs(model_list1.scl, oakplots.sub,
          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))

# sem.coefs(oak_and_bay_list1, oak_and_bay_plots, standardize = "none",
#           corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log",
#                           "tot_bay.log ~~ H.2005"))
```

```{r path models using avg dry season tmax}
# Model set using average daily maximum temperature during the dry season.
model_avgtmax_ds <- list(
      shannons <- lm(H.2005 ~ twi15m.log, data = oakplots.sub),
      # umca_density <- lm(tot_bay.log ~ twi15m.log, data = oakplots.sub),

      slc <- lmer(tot_lfct.log ~ tot_bay.log + avg_tmax_ds + avg_hrs1422_wet_t_t1.log + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),

      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ tot_bay.log + (1|plotid) + (1|sample_year), data = oakplots.sub),

      avgtmax_ds <- lmer(avg_tmax_ds ~ tot_bay.log + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),

      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + avg_tmax_ds + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.sub)
)

model_avgtmax_ds.scl <- list(
      shannons <- lm(shannons2005.scl ~ twi15m.log.scl + (1|sample_year), data = oakplots.sub),
      # umca_density <- lm(tot_bay.log ~ twi15m.log, data = oakplots.sub),

      slc <- lmer(tot_lfct.log.scl ~ tot_bay.log.scl + avg_tmax_ds.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),

      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log.scl ~ tot_bay.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),

      avgtmax_ds <- lmer(avg_tmax_ds.scl ~ tot_bay.log.scl + twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),

      oak_infection <- glmer(oak.inf ~ tot_lfct.log.scl + avg_hrs1422_wet_t_t1.log.scl + avg_tmax_ds.scl + shannons2005.scl + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.sub)
)

oak_and_bay_avgtmax_ds <- list(
      shannons <- lm(H.2005 ~ twi15m.log, data = filter(oak_and_bay_plots)),
      # umca_density <- lm(tot_bay.log ~ twi15m.log, data = oak_and_bay_plots),

      slc <- lmer(tot_lfct.log ~ tot_bay.log + avg_tmax_ds + avg_hrs1422_wet_t_t1.log + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oak_and_bay_plots),

      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ tot_bay.log + (1|plotid) + (1|sample_year), data = oak_and_bay_plots),

      avgtmax_ds <- lmer(avg_tmax_ds ~ tot_bay.log + twi15m.log + (1|plotid) + (1|sample_year), data = oak_and_bay_plots),

      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + avg_tmax_ds + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oak_and_bay_plots)
)

sem.fit(model_avgtmax_ds, oakplots.sub,
        corr.errors = c(
              "avg_tmax_ds ~~ avg_hrs1422_wet_t_t1.log",
              "H.2005 ~~ avg_tmax_ds"
                        )
        )
sem.fit(oak_and_bay_avgtmax_ds, oak_and_bay_plots,
        corr.errors = c("avg_tmax_ds ~~ avg_hrs1422_wet_t_t1.log"
                        ,"H.2005 ~~ tot_bay.log"
                        # ,"H.2005 ~~ avg_tmax_ds"
                        )
        )

sem.coefs(model_avgtmax_ds, oakplots.sub, standardize = "none",
          corr.errors = c(
                "avg_tmax_ds ~~ avg_hrs1422_wet_t_t1.log",
                "H.2005 ~~ avg_tmax_ds"))
sem.coefs(model_avgtmax_ds.scl, oakplots.sub, standardize = "none",
          corr.errors = c(
                "avg_tmax_ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl",
                "shannons2005.scl ~~ avg_tmax_ds.scl"))

sem.coefs(oak_and_bay_avgtmax_ds, oak_and_bay_plots, standardize = "none",
          corr.errors = c("avg_tmax_ds ~~ avg_hrs1422_wet_t_t1.log",
                        "H.2005 ~~ tot_bay.log"))
```

Exploratory modeling of symptomatic leaf count with different temperature variables
```{r slc with dry-season temperature variables}
# Models using dry-season temperatures (June-September)
slc_abv25ds <- lmer(tot_lfct.log ~ tot_bay.log + hrs_abv25ds + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

slc_abv20ds <- lmer(tot_lfct.log ~ tot_bay.log + hrs_abv20ds + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

slc_blw18ds <- lmer(tot_lfct.log ~ tot_bay.log + hrs_blw18ds + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

slc_avgtmax_ds <- lmer(tot_lfct.log ~ tot_bay.log + avg_tmax_ds + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

slc_avgtmin_ds <- lmer(tot_lfct.log ~ tot_bay.log + avg_tmin_ds + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

AIC(slc_avgtmax_ds)
AIC(slc_avgtmin_ds)
AIC(slc_abv25ds)
AIC(slc_abv20ds)
AIC(slc_blw18ds)
## The best model based on AIC is avgtmax_ds. Interestingly, abv20ds & blw18ds have precisely equal support, but show opposite inference. More hours above 20C during the dry season reduces symptomatic leaf count the following spring, while more hours below 18C increases symptomatic leaf count.
library(ggplot2)
par(mfrow = c(2,1))
boxplot(tot_lfct.log ~ sample_year, data = oakplots.sub,
        xlab = "sample year", ylab = "ln(total symptomatic leaf count)")
boxplot(hrs_abv25ds ~ sample_year, data = oakplots.sub,
        xlab = "sample_year", ylab = "number of dry-season hours >25C")
boxplot(avg_lfct.log ~ sample_year, data = oakplots.sub,
        xlab = "sample year", ylab = "ln(total symptomatic leaf count)")

ggplot(oak_and_bay_plots, aes(hrs_abv25ds, tot_lfct.log)) +
      geom_point() +
      ggtitle("Oak & Bay laurel Plots") +
      xlab("Dry-season Hours >25C") + ylab("log(Total Symptomatic Leaf Count")
ggplot(oakplots.sub, aes(hrs_abv25ds, tot_lfct.log)) +
      geom_point() +
      ggtitle("Oak Plots") +
      xlab("Dry-season Hours >25C") + ylab("log(Total Symptomatic Leaf Count")

ggplot(oakplots.sub, aes(hrs_abv20ds, tot_lfct.log)) +
      geom_point() +
      ggtitle("Oak Plots") +
      xlab("Dry-season Hours >20C") + ylab("log(Total Symptomatic Leaf Count")
ggplot(oakplots.sub, aes(hrs_blw18ds, tot_lfct.log)) +
      geom_point() +
      ggtitle("Oak Plots") +
      xlab("Dry-season Hours <18C") + ylab("log(Total Symptomatic Leaf Count")

ggplot(oakplots.sub, aes(avg_tmax_ds, tot_lfct.log)) +
      geom_point() +
      ggtitle("Oak Plots") +
      xlab("Dry-season Average Daily Maximum Temperature (C)") + 
      ylab("log(Total Symptomatic Leaf Count")

ggplot(oakplots.sub, aes(avg_tmin_ds, tot_lfct.log)) +
      geom_point() +
      ggtitle("Oak Plots") +
      xlab("Dry-season Average Daily Minimum Temperature (C)") + 
      ylab("log(Total Symptomatic Leaf Count")
```

## Models of total symptomatic leaf-count using rainy season temperatures
These models use different rainy-season (November-May) temperature variables to model total symptomatic leaf count on California bay laurel
```{r slc models using rainy season temperature}
# Models using rainy-season temperatures (November-May)
slc_avgtmax_rs <- lmer(tot_lfct.log ~ tot_bay.log + avg_tmax_rs + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

slc_avgtmin_rs <- lmer(tot_lfct.log ~ tot_bay.log + avg_tmin_rs + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

slc_hrs1420_rs <- lmer(tot_lfct.log ~ tot_bay.log + hrs_14_20_rs + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

slc_hrs_abv25rs <- lmer(tot_lfct.log ~ tot_bay.log + hrs_abv25rs + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

slc_hrs_abv20_rs <- lmer(tot_lfct.log ~ tot_bay.log + hrs_abv20_rs + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

slc_hrs_blw10rs <- lmer(tot_lfct.log ~ tot_bay.log + hrs_blw10rs + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

AIC(slc_avgtmax_rs)
AIC(slc_avgtmin_rs)
AIC(slc_hrs1420_rs)
AIC(slc_hrs_abv25rs)
AIC(slc_hrs_abv20_rs)
AIC(slc_hrs_blw10rs)
```

## Models of symptomatic leaf count using temperature on wet days
These models use different temperature variables calculated for days that had measurable rainfall during the rainy season, based on plots being assigned to the nearest rain gauge.
```{r slc models using wet-days temperatures}
# Models using wet-days temperatures
slc_hrsblw14_wet <- lmer(tot_lfct.log ~ tot_bay.log + hrsblw14_wet + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

slc_hrsabv22_wet <- lmer(tot_lfct.log ~ tot_bay.log + hrsabv22_wet + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

slc_hrs1422_wet <- lmer(tot_lfct.log ~ tot_bay.log + hrs1422_wet + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

slc_hrs1020_wet <- lmer(tot_lfct.log ~ tot_bay.log + hrs1020_wet + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

slc_hrsabv20_wet <- lmer(tot_lfct.log ~ tot_bay.log + hrsabv20_wet + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

slc_hrsblw10_wet <- lmer(tot_lfct.log ~ tot_bay.log + hrsblw10_wet + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

slc_avgtmin_wet <- lmer(tot_lfct.log ~ tot_bay.log + avgtmin_wet + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

slc_avgtmax_wet <- lmer(tot_lfct.log ~ tot_bay.log + avgtmax_wet + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

slc_avgtemp_wet <- lmer(tot_lfct.log ~ tot_bay.log + avgtemp_wet + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

AIC(slc_hrsblw14_wet)
AIC(slc_hrs1422_wet)
AIC(slc_hrsabv22_wet)
AIC(slc_hrsblw10_wet)
AIC(slc_hrs1020_wet)
AIC(slc_hrsabv20_wet)
AIC(slc_avgtmax_wet)
AIC(slc_avgtmin_wet)
AIC(slc_avgtemp_wet)
```

## Models using lagged dry-season temperature variables

```{r slc models using lagged dry-season temperature}
# Models using lagged dry-season temperature variables
slc_hrs_abv25ds_tminus1 <- lmer(tot_lfct.log ~ tot_bay.log + hrs_abv25ds_tminus1 + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

slc_hrs_blw18ds_tminus1 <- lmer(tot_lfct.log ~ tot_bay.log + hrs_blw18ds_tminus1 + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

slc_avg_tmax_ds_tminus1 <- lmer(tot_lfct.log ~ tot_bay.log + avg_tmax_ds_tminus1 + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

slc_avg_hrsblw18_ds_tminus1 <- lmer(tot_lfct.log ~ tot_bay.log + avg_hrsblw18_ds_tminus1 + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

slc_avgtmax_ds_t_t1 <- lmer(tot_lfct.log ~ tot_bay.log + avgtmax_ds_t_t1 + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

slc_avg_hrsabv25_ds_tminus1 <- lmer(tot_lfct.log ~ tot_bay.log + avg_hrsabv25_ds_tminus1 + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)

AIC(slc_hrs_abv25ds_tminus1)
AIC(slc_hrs_blw18ds_tminus1)
AIC(slc_avg_tmax_ds_tminus1)
AIC(slc_avg_hrsblw18_ds_tminus1)
AIC(slc_avg_hrsabv25_ds_tminus1)
AIC(slc_avgtmax_ds_t_t1)
```