### 2019-04-07 pathogen-spillover-diss-ch.Rmd
Old update?

---
title: "Pathogen Spillover - dissertation chapter"
author: "Whalen Dillon"
date: "July 18, 2016"
output:
  html_document:
    self_contained: no
    toc: yes
  pdf_document: default
  word_document: default
---

# Appendix 2: Analysis of Environmental Controls on Pathogen Spillover
**Effects of Diversity, Topography, and Interannual Climate Variability**

[Conceptual Path Model](figures/conceptual-path-model-updated.png)

## Results

```{r set options, echo=F}
# knitr::opts_chunk$set(fig.width = 7.5, fig.height = 7)
knitr::opts_chunk$set(dpi = 300)
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)
```

```{r load data, echo=F, message=F}
library(knitr)
load("climate_path_model_data.RData")
library(piecewiseSEM)# version 1.1.1, tried most recent version & kept getting errror 7/18/2016 b/c matrices rows were unequal in sem.missing.paths function
## Avoided this error in 1.1.2 by removing observations with NA values, which gives the same results as before, so maybe it wasn't properly dropping levels when extracting info from the models.
library(plyr); library(dplyr, verbose = F)
library(lme4, verbose = F); library(nlme, verbose = F)
# library(lmerTest, verbose = F)

panel.cor <- function(x, y, digits = 2, cex.cor, ...)
{
      usr <- par("usr"); on.exit(par(usr))
      par(usr = c(0, 1, 0, 1))
      # correlation coefficient
      r <- cor(x, y)
      txt <- format(c(r, 0.123456789), digits = digits)[1]
      txt <- paste("r= ", txt, sep = "")
      text(0.5, 0.6, txt)

      # p-value calculation
      # p <- cor.test(x, y)$p.value
      # txt2 <- format(c(p, 0.123456789), digits = digits)[1]
      # txt2 <- paste("p= ", txt2, sep = "")
      # if(p<0.01) txt2 <- paste("p= ", "<0.01", sep = "")
      # text(0.5, 0.4, txt2)
}

# Calculate oak density variable as stems per hectare 
oakplots.sub$umca_density_ha <- (oakplots.sub$tot_bay / (15*15))*10000
oakplots.sub$umca_density_ha.log <- log(oakplots.sub$umca_density_ha+1)
```

```{r rescale variables, echo=F}
## rescale variables
# oakplots.sub2$inf_bay.scl <- scale((oakplots.sub2$infected_bay_ct))
oakplots.sub$tot_bay.scl <- as.numeric(scale(oakplots.sub$tot_bay))
oakplots.sub$tot_bay.log.scl <- as.numeric(scale(oakplots.sub$tot_bay.log))
oakplots.sub$umca_density_ha.scl <- as.numeric(scale(oakplots.sub$umca_density_ha))
oakplots.sub$umca_density_ha.log.scl <- as.numeric(scale(oakplots.sub$umca_density_ha.log))
oakplots.sub$umca_basal_area.scl <- as.numeric(scale(oakplots.sub$umca_basal_area))
oakplots.sub$umca_ba.log.scl <- as.numeric(scale(oakplots.sub$umca_ba.log))
oakplots.sub$tot_lfct.scl <- as.numeric(scale(oakplots.sub$tot_lfct))
oakplots.sub$tot_lfct.log.scl <- as.numeric(scale(oakplots.sub$tot_lfct.log))
# oakplots.sub$avg_lfct.scl <- scale(oakplots.sub$avg_lfct)
oakplots.sub$hrs_abv25ds.scl <- as.numeric(scale(oakplots.sub$hrs_abv25ds))
oakplots.sub$dys_abv25ds.scl <- as.numeric(scale(oakplots.sub$dys_abv25ds))
oakplots.sub$avg_tmax_ds.scl <- as.numeric(scale(oakplots.sub$avg_tmax_ds))
oakplots.sub$avg_hrsblw14_wet_tminus1.scl <- as.numeric(scale(oakplots.sub$avg_hrsblw14_wet_tminus1))
oakplots.sub$hrs1422_wet_tminus1.log.scl <- as.numeric(scale(oakplots.sub$hrs1422_wet_tminus1.log))
# oakplots.sub$avg_hrs1422_wet_t_t1.scl <- scale(oakplots.sub$avg_hrs1422_wet_t_t1)
oakplots.sub$avg_hrs1422_wet_t_t1.log.scl <- as.numeric(scale(oakplots.sub$avg_hrs1422_wet_t_t1.log))

## static variables
oakplots.sub$candens15m.scl <- as.numeric(scale(oakplots.sub$candens15m))
oakplots.sub$shannons2005.scl <- as.numeric(scale(oakplots.sub$H.2005))
oakplots.sub$shannons2014.scl <- as.numeric(scale(oakplots.sub$H.2014))
oakplots.sub$twi15m.scl <- as.numeric(scale(oakplots.sub$twi15m))
oakplots.sub$twi15m.log.scl <- as.numeric(scale(oakplots.sub$twi15m.log))

## make consolidated data frame only of variables in the path model
oakplots.model.data <- select(oakplots.sub, plotid, sample_year, oak.inf, tot_lfct, tot_lfct.log, hrs_abv25ds, avg_hrs1422_wet_t_t1, avg_hrs1422_wet_t_t1.log, umca_density_ha, umca_density_ha.log, H.2005, twi15m, twi15m.log#, twi15m.log.scl, tot_lfct.log.scl, hrs_abv25ds.scl, avg_hrs1422_wet_t_t1.log.scl, umca_density_ha.log.scl, shannons2005.scl
                              )

oakplots.model.data <- filter(oakplots.model.data, avg_hrs1422_wet_t_t1 !="NA", H.2005 != "NA", tot_lfct != "NA")
write.csv(oakplots.model.data, "analysis/data/oakplots-pathmodel-data.csv",
          row.names = F)
```

### Path model 
This is the list of models that make up the path model to be fit using the methods implemented in the 'piecewiseSEM' packag. The continuous variables in the models were standardized using the default settings of the 'scale' function in R, resulting in each having a mean of zero and standard deviation of one. Each variable is first centered by subtracting its mean and then scaled by dividing by its standard deviation.

```{r hours >25 path model no umca density unscaled, echo=F, eval=F}
model_list2 <- list(
      shannons <- lm(H.2005 ~ twi15m.log, data = filter(oakplots.model.data, sample_year == 2006)),

      # umca_density <- lmer(umca_density_ha.log ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),

      slc <- lmer(tot_lfct.log ~ hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      hrs_abv25 <- lmer(hrs_abv25ds ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.model.data)
)
```

```{r hours >25 path model no umca density scaled, echo=F, eval=F}
model_list2.scl <- list(
      shannons <- lm(shannons2005.scl ~ twi15m.log.scl, data = filter(oakplots.model.data, sample_year == 2006)),

      # umca_density <- lmer(umca_density_ha.log.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      slc <- lmer(tot_lfct.log.scl ~ hrs_abv25ds.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      hrs_abv25 <- lmer(hrs_abv25ds.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      oak_infection <- glmer(oak.inf ~ tot_lfct.log.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.model.data)
)
```

```{r hours >25 path model all oak plots unscaled, echo=F}
# slc_nlme <- lme(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005, random = ~1|plotid, data = oakplots.model.data)

model_list1 <- list(
      shannons <- lm(H.2005 ~ twi15m, data = filter(oakplots.model.data, sample_year == 2006)),

      umca_density <- lmer(umca_density_ha.log ~ twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      slc <- lmer(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ 
                                   # umca_density_ha.log + twi15m.log +
                                   twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      hrs_abv25 <- lmer(hrs_abv25ds ~ 
                              # umca_density_ha.log + 
                              twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.model.data)
)

model_list1_refit <- list(
      shannons_refit <- lm(H.2005 ~ twi15m, data = filter(oakplots.model.data, sample_year == 2006)),

      umca_density_refit <- lmer(umca_density_ha.log ~ H.2005 + twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      slc_refit <- lmer(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds +
                        #avg_hrs1422_wet_t_t1.log + H.2005 + 
                        (1|plotid) + (1|sample_year), 
                  data = oakplots.model.data),

      wet_hours_1422_refit <- lmer(avg_hrs1422_wet_t_t1.log ~ 
                                   # umca_density_ha.log + twi15m.log +
                                   twi15m +
                                   (1|plotid) + (1|sample_year), 
                             data = oakplots.model.data),

      hrs_abv25_refit <- lmer(hrs_abv25ds ~ 
                              # umca_density_ha.log + 
                              twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      oak_infection_refit <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.model.data)
)
```

```{r hours >25 path model all oak plots scaled, eval = F}
model_list1.scl <- list(
      shannons <- lm(shannons2005.scl ~ twi15m.log.scl, data = filter(oakplots.model.data, sample_year == 2006)),

      umca_density <- lmer(umca_density_ha.log.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      slc <- lmer(tot_lfct.log.scl ~ umca_density_ha.log.scl + hrs_abv25ds.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log.scl ~
                                   # umca_density_ha.log.scl + 
                                   twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      hrs_abv25 <- lmer(hrs_abv25ds.scl ~ 
                              # umca_density_ha.log.scl + 
                              twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      oak_infection <- glmer(oak.inf ~ tot_lfct.log.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.model.data)
)

## remove microclimate variables
model_list1.scla <- list(
      shannons <- lm(shannons2005.scl ~ twi15m.log.scl, data = filter(oakplots.model.data, sample_year == 2006)),

      umca_density <- lmer(umca_density_ha.log.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      slc <- lmer(tot_lfct.log.scl ~ umca_density_ha.log.scl + shannons2005.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      # wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log.scl ~
      #                              # umca_density_ha.log.scl + 
      #                              twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),
      # 
      # hrs_abv25 <- lmer(hrs_abv25ds.scl ~ 
      #                         # umca_density_ha.log.scl + 
      #                         twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),

      oak_infection <- glmer(oak.inf ~ tot_lfct.log.scl + shannons2005.scl + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.model.data)
)
```

In my assessment of model residuals I found no substantial violations of the assumption of homoscedacity for individual models (although there was some), and no relationships between the model residuals and the predictor variable(s). The patterning in the residual plots for some of the models is likely in part due to observations of count variables where the value is zero. Since the dataset in this analysis is all plots with one or more susceptible oak species there are some plots that have zero bay laurel and zero symptomatic bay laurel leaves.

### Examining the residual plots of each piece of the path model

**Residual plots of Shannon's Diversity Index model**
```{r shannons residuals plot, echo=F}
plot(shannons, main = "OLS regression of diversity ~ wetness index residuals")
plot(shannons$model$twi15m, resid(shannons),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "Diversity model residuals vs. wetness index predictor")
```

**Residual plots of bay laurel density model**
```{r bay laurel density residuals plot, echo=F}
plot(umca_density, main = "LMM of bay laurel density residuals")
plot(umca_density@frame$twi15m, resid(umca_density),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "Bay laurel density model residuals vs. wetness index predictor")
```

**Residual plots of symptomatic leaf count model**
```{r symptomatic leaf count residuals plot, echo=F}
plot(resid(slc), main = "LMM of symptomatic leaf count residuals")
# qqnorm(resid(slc), main="Q-Q plot for conditional residuals - SLC")
# qqnorm(ranef(slc)$sample_year$`(Intercept)`, 
#        main="Q-Q plot for Sample Year")
# qqnorm(ranef(slc)$plotid$`(Intercept)`, 
#        main="Q-Q plot for plot ID")

plot(slc@frame$umca_density_ha.log, resid(slc),
     xlab = "bay laurel density",
     ylab = "model residuals",
     main = "SLC model residuals vs. bay laurel density predictor")
plot(slc@frame$hrs_abv25ds, resid(slc),
     xlab = "# dry-season hours >25 C",
     ylab = "model residuals",
     main = "SLC model residuals vs. dry-season hours >25 C predictor")
plot(slc@frame$avg_hrs1422_wet_t_t1.log,
     xlab = "avg. wet-hours 14-22C",
     ylab = "model residuals",
     main = "SLC model residuals vs. wet-hours predictor")
plot(slc@frame$H.2005, resid(slc),
     xlab = "Shannon's diversity index",
     ylab = "model residuals",
     main = "SLC model residuals vs. Shannon's diversity predictor")
# plot(slc@frame$twi15m.log.scl, resid(slc),
#      xlab = "wetness index",
#      ylab = "model residuals",
#      main = "SLC model residuals vs. wetness index predictor")
```
```{r pairs plot of slc predictor variables, echo=F}
pairs(na.omit(select(oakplots.sub, umca_density_ha.log, hrs_abv25ds, avg_hrs1422_wet_t_t1.log, H.2005, twi15m)),
      lower.panel = panel.cor, diag.panel = panel.hist,
      main = "Predictor variables for symptomatic leaf count models")
```

**Residual plots of average number of wet-hours model**
```{r wet-hours residuals plot, echo=F}
plot(wet_hours_1422, main = "LMM of avg. # of hours 14-22 C current & prev. year residuals")
# plot(wet_hours_1422@frame$umca_density_ha.log.scl, resid(wet_hours_1422),
#      xlab = "bay laurel density",
#      ylab = "model residuals",
#      main = "Wet-hours model residuals vs. bay laurel density predictor")
plot(wet_hours_1422@frame$twi15m, resid(wet_hours_1422),
     xlab = "bay laurel density",
     ylab = "model residuals",
     main = "Wet-hours model residuals vs. bay laurel density predictor")
```

**Residual plots of dry-season temperature >25C model**
```{r dry-season hours residuals plot, echo=F}
plot(hrs_abv25, main = "LMM of # of hours >25 C during prev. dry season residuals")
plot(hrs_abv25@frame$twi15m, resid(hrs_abv25),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "Dry-season hours model residuals vs. wetness index predictor")
# plot(hrs_abv25@frame$umca_density_ha.log.scl, resid(hrs_abv25),
#      xlab = "wetness index",
#      ylab = "model residuals",
#      main = "Dry-season hours model residuals vs. wetness index predictor")
```

**Residual plots of oak infection model**
```{r oak infection residuals plot, echo=F}
plot(oak_infection, main = "Binomial GLMM of oak infection residuals")
plot(oak_infection@frame$tot_lfct.log, resid(oak_infection),
     xlab = "symptomatic leaf count",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. symptomatic leaf count predictor")
plot(oak_infection@frame$avg_hrs1422_wet_t_t1.log, resid(oak_infection),
     xlab = "avg. wet-hours 14-22C",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. wet-hours predictor")
plot(oak_infection@frame$H.2005, resid(oak_infection),
     xlab = "Shannon's diversity index",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. diversity predictor")
```
```{r pairs plot of oak infection predictor variables, echo=F}
pairs(na.omit(select(oakplots.sub, tot_lfct.log, avg_hrs1422_wet_t_t1.log, H.2005)),
      lower.panel = panel.cor, diag.panel = panel.hist,
      main = "Predictor variables for oak infection model")
```

### Fitting the path model

I fit the path model by implementing the d-sep test using the 'sem.fit' function separately for standardized and unstandardized variables, with the assumption of correlated errors between the two climate variables. In the model with standardized variables all continuous variables were scaled to a mean of zero and standard deviation of one. The results for the conditional claims of independence (the d-sep test) for the standardized and unstandardized variables are reported in the tables. Statistically significant p-values indicate possible missing paths, however, the overall model may still be accepted if the p-value of Fisher's C-statistic is >0.05. The results of the d-sep test assessing overall model fit is reported with Fisher's C-statistic (fisher.c), degrees of freedom (df), the p-value of Fisher's C based on comparison to a Chi-squared distribution, the general Akaike information criterion (AIC) and corrected for sample size (AICc), the number of parameters (K) estimated, the sample size (n). The sample size and number of parameters are used in calculating AIC.

In this case, the significant missing path between diversity and bay laurel density may be explained ecologically, but neither variable can be determined as causal to the other. I chose to maintain the initial hypothesized structure of the path model since the overall fit of the model was still acceptable, though this relationship could also be included as a correlated error.

```{r fit and compare all oak plots, echo=F, message=F, warning=F}
raw_variables_model <- sem.fit(model_list1, oakplots.model.data, 
        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
        conditional = T, .progressBar = F)

raw_vars_refit <- sem.fit(model_list1_refit, oakplots.model.data, 
        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
        conditional = T, .progressBar = F)

# scaled_variables_model <- sem.fit(model_list1.scl, oakplots.model.data,
#           corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"),
#           conditional = T, .progressBar = F)
# scaled_variables_modela <- sem.fit(model_list1.scla, oakplots.model.data,
#           # corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"),
#           conditional = F, .progressBar = F)

# raw_variables_model_noUMCA <- sem.fit(model_list2, oakplots.sub, 
#         corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
#         .progressBar = F)
# scaled_variables_model_noUMCA <- sem.fit(model_list2.scl, oakplots.sub,
#           corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"),
#           .progressBar = F)
```

```{r tables of d-sep test results from sem.fit, echo=F}
kable(raw_variables_model, digits = 4,
      caption = "Results of the d-sep test using raw variables.")
# kable(scaled_variables_model,
#       caption = "Results of the d-sep test using standardized variables.")

# kable(raw_variables_model_noUMCA, 
#       caption = "Results of the d-sep test using raw variables - no UMCA density.")
# kable(scaled_variables_model_noUMCA,
#       caption = "Results of the d-sep test using standardized variables - no UMCA density.")
```

### Path coefficients
The path model provides direct and indirect effects of predictor variables on multiple responses, which can be combined to calculate the total effect of a predictor variable on a response. Direct effects are from the variables that have a path pointing directly to a response variable, so effect on the response is the value of the path coefficient. Variables with indirect effects are those that are at least once removed from another response variable, that is the response variable of a predictor with an indirect effect is the predictor variable along another path in the model. Indirect effects are calculated by multiplying the coefficients along the path to the response variable of interest. The total effect of variables that have a direct and indirect effect(s) on a response is the sum of the direct and indirect effects. The standardized coefficients of the predictors that have the same response variable (i.e. direct effects) can be interpreted relative to each other in terms of influence on the response variable.

```{r get path coefficients, echo=F}
raw_coefficients <- sem.coefs(model_list1, oakplots.model.data,
          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
scaled_coefficients <- sem.coefs(model_list1, oakplots.model.data,
          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
          standardize = "scale")

raw_coeff_refit <- sem.coefs(model_list1_refit, oakplots.model.data,
          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
scaled_coeff_refit <- sem.coefs(model_list1_refit, oakplots.model.data,
          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
          standardize = "scale")

# raw_coefficients_noUMCA <- sem.coefs(model_list2, oakplots.sub,
#           corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
# scaled_coefficients_noUMCA <- sem.coefs(model_list2.scl, oakplots.sub,
#           corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))
```

**Tables of the path coeffcients for the fitted path models with raw and standardized variables.**

```{r tables of path coefficients, echo=F}
kable(raw_coefficients, digits = 4,
      caption = "Raw path coefficients")
# kable(raw_coefficients_noUMCA,
#       caption = "Raw path coefficients - no UMCA density")

kable(scaled_coefficients, digits = 4,
      caption = "Standardized path coefficients")
# kable(scaled_coefficients_noUMCA,
#       caption = "Standardized path coefficients - no UMCA density")
```

There were five non-significant path coefficients at the p=0.05 level. Two of these (symptomatic leaf count ~ wet-hours and wet-hours ~ bay laurel density) were marginal, with p-values <0.10, which may warrant further investigation and I will address the ecological relevance in the Discussion section. The topographic wetness index was not a statistically significant predictor of bay laurel density nor symptomatic leaf count after accounting for other predictor variables and additional variability due to unmeasured plot-level factors and repeated measures. Locations with higher potential soil moisture (larger values of the wetness index) were significantly less diverse and had significanly fewer hours when the temperature was above 25 C during the dry-season. Based on the path model structure, the wetness index has only indirect effects on oak infection (pathogen spillover). The inference for the total effect of the wetness index on pathogen spillover was positive, indicating that oak trees located in areas with higher potential soil moisture are more likely to become infected.

The diversity index was not a statistically significant predictor of symptomatic leaf count, but it was the strongest of the direct effects on oak infection. The negative effect of diversity on oak infection is evidence for a dilution effect, where oak trees in plots with greater plant diversity are less likely to become infected. With the other hypothesized relationship of diversity in this system non-significant I only interpret the direct effect.

The average number of hours between 14 C and 22 C on wet days during the current and sample year had a positive, but statistically weak direct effect (p = 0.068) on symptomatic leaf count. The direct effect of this variable on oak infection was positive and statistically significant, resulting in an overall positive effect. This indicates that oak trees in locations that have warm temperatures when it is wet are more likely to become infected.

The number of hours greater than 25 C during the dry-season prior to the sample year was hypothesized to have a direct effect on bay laurel symptomatic leaf count (potential inoculum), and through this an indirect effect on pathogen spillover leading to oak infection. The direct effect on potential inoculum was negative, indicating that bay laurel trees in locations experiencing more time at desiccating temperatures during the dry-season had fewer infected leaves the during the following sample season. The total effect of the dry-season hours on pathogen spillover (oak infection) is negative, so oak trees in locations that experienced more time at desiccating temperatures are less likely to become infected.

Expectedly, bay laurel density and potential inoculum load are tightly related through the direct effect of density on symptomatic leaf count. I structured the path model hypothesizing that bay laurel density indirectly influences pathogen spillover through its direct effect on inoculum load. The total effect of bay laurel density effect on pathogen spillover is positive through its direct influences on inoculum load and average number of wet-hours.

The symptomatic leaf count represents potential inoculum load, and ecologically is the primary driver that I measured of the pathogen spillover process leading to oak infection. Oaks in locations with greater inoculum load are more likely to be infected. Additional factors have direct and indirect effects on the success of the pathogen spillover process as indicated by infection of oak trees.

By calculating oak infection at the plot-level we aggregated any variation in susecptibility that may be present in different species or individuals.

## Disease Figures
```{r disease figures, echo=F, warning=F}
library(ggplot2)
temperatures <- left_join(
      select(summer_temps, plotid, sample_year, hrs_abv25ds),
      select(wet_hrs, plotid, sample_year, hrs1422_wet),
      by = c("plotid","sample_year"))
# oak_sod$year <- as.Date(oak_sod$sample_year)
temps_figure <- ggplot(oak_sod, aes(group = sample_year))

(oak_infection_figure <- temps_figure +
      geom_bar(aes(as.factor(sample_year), inf_oak_ct), stat = "identity") +
      # geom_boxplot(aes(as.factor(sample_year), inf_oak_ct)) +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Number of infected oaks") +
      ggtitle("Oak infection")
)

(slc_avg_figure <- temps_figure + 
      geom_boxplot(aes(as.factor(sample_year), avg_lfct), notch = F) +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Pathogen Load \n(average symptomatic leaf count)") +
      ggtitle("California bay laurel leaf symptoms - plot averages")
)

(slc_total_figure <- temps_figure + 
      geom_boxplot(aes(as.factor(sample_year), tot_lfct), notch = F) +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Pathogen Load \n(total symptomatic leaf count)") +
      ggtitle("California bay laurel leaf symptoms - plot totals")
)
```

## Temperature Figures
```{r temperature figures, echo=F, warning=F, eval=T}
library(ggplot2)
library(scales)
temperatures <- left_join(
      select(summer_temps, plotid, sample_year, hrs_abv25ds),
      select(wet_hrs, plotid, sample_year, hrs1422_wet),
      by = c("plotid","sample_year"))
# oak_sod$year <- as.Date(oak_sod$sample_year)

temps_figure <- ggplot(oak_sod, aes(group = sample_year))

(sstemps_figure <- temps_figure + 
      geom_boxplot(aes(as.factor(sample_year), hrs_abv25ds), notch = T) + 
      geom_hline(aes(yintercept = mean(hrs_abv25ds)), colour = "red") +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Number of hours") +
      ggtitle("Heat Exposure (dry-season Hours >25 C)")
)

(rstemps_figure <- temps_figure + 
      geom_boxplot(aes(as.factor(sample_year), hrs1422_wet), notch = T) +
      geom_hline(aes(yintercept = mean(hrs1422_wet)), colour = "red") +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Number of hours") +
      ggtitle("Warm & Wet Conditions (wet-hours 14-22 C)")
)
```

## Rainfall Figure
Total annual rainfall estimated at plot locations using two-dimensional interpolation methods implemented in GRASS-GIS. The red line indicates the average rainfall from 2005-2014.

```{r plot rainfall figure, echo=F}
library(ggplot2)
# summary(rainy_season_ppt)
# rainy_season_ppt$year <- as.factor(rainy_season_ppt$year)
rainfall_figure <- ggplot(filter(rainy_season_ppt, year > 2004), 
                          aes(year, rain2d_total))
rainfall_figure +
      geom_boxplot(aes(group = year), notch = T) +
      geom_hline(aes(yintercept = mean(rain2d_total)),
                 colour = "red") +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Total Rainfall (mm)") +
      ggtitle("Rainfall Estimated at Plots from Rain Gauges (November - May)")

rainy_days_figure <- ggplot(filter(rainy_season_ppt, year > 2004), 
                          aes(year, rain_days_v))
rainy_days_figure +
      geom_boxplot(aes(group = year), notch = F) +
      geom_hline(aes(yintercept = mean(rain_days_v)),
                 colour = "red") +
      theme_bw() +
      xlab("Sample Year") +
      ylab("Number of days w/precipitation") +
      ggtitle("Rainy Days at Plots (November - May)")
```

## Session Information
```{r print session info, echo=FALSE}
print(sessionInfo())
```

## Model a subset of the data
I want to examine the different matrices of the mixed effects models using a smaller datset that will make them easier to visualize.
```{r model subset data, eval=F}
small_data <- oakplots.model.data[1:103,]
xtabs(~sample_year + plotid, droplevels(small_data))
## shows a somewhat unbalanced data set

small_data_model_list <- list(
      shannons_small <- lm(H.2005 ~ twi15m, data = filter(small_data, sample_year == 2006)),# this model avoids pseudoreplication

      umca_density_small <- lmer(umca_density_ha.log ~ twi15m + (1|plotid) + (1|sample_year), data = small_data),

      slc_small <- lmer(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), data = small_data),

      wet_hours_1422_small <- lmer(avg_hrs1422_wet_t_t1.log ~ 
                                   # umca_density_ha.log + twi15m.log +
                                   twi15m + (1|plotid) + (1|sample_year), data = small_data),

      hrs_abv25_small <- lmer(hrs_abv25ds ~ 
                              # umca_density_ha.log + 
                              twi15m + (1|plotid) + (1|sample_year), data = small_data),

      oak_infection_small <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = small_data)
)

summary(umca_density_small)
summary(slc_small)
summary(hrs_abv25_small)
anova(umca_density,update(umca_density, .~ 1+. -twi15m))

getME(umca_density_small, "X")# fixed-effects model matrix
getME(umca_density_small, "Z")# random-effects model matrix
getME(umca_density_small, "Zt")# transposed random-effects model matrix, diagonal values = 1
# image(getME(umca_density_small, "Z"))#
getME(umca_density_small, "Ztlist")# list of components of Zt separated by individual variance component
getME(umca_density_small, "mmList")# list of raw model matrices associated w/random effects terms
getME(umca_density_small, "u")# conditional mode of "spherical" random effects variable
getME(umca_density_small, "b")# conditional mode of random effects variable
getME(umca_density_small, "Gp")

image(getME(umca_density_small, "L"))# sparse Cholesky factor matrix of the penalized random-effects model
image(getME(umca_density_small, "Lambda"))# relative covariance factor of the random effects
image(getME(umca_density_small, "Lambdat"))# cross-product of random effects model matrix, diagonal values = random-effects parameter estimates

getME(umca_density_small, "A")# scaled sparse model matrix
getME(umca_density_small, "RX")# Cholesky factor for fixed-effects parameters
getME(umca_density_small, "RZX")#cross-term in the full Cholesky factor

getME(umca_density_small, "sigma")# residual standard error

getME(umca_density_small, "theta")# random-effects parameter estimates

getME(umca_density_small, "q")# number of coluns of RE model matrix Z
getME(umca_density_small, "l_i")# number of levels of grouping factors
getME(umca_density_small, "k")# number of RE terms
getME(umca_density_small, "m_i")# number of covariance parameters in each RE term
getME(umca_density_small, "devcomp")# list of items describing fitted model

# image(getME(oak_infection, "L"))
```